---
title: "Eelgrass_density_2018"
author: "Lia Domke"
date: "5/28/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Preparing data for archival in KNB. Data has been entered and QA/QC'd and saved in csv format. These data are from field season 2018 and include data from 8 sites across the western coast of Prince of Wales Island, AK. 

At two sites in 2018, a 120 m transect was used with 12 0.25 m2 quadrats placed approximately every 10 m within the seagrass meadows (sites: Photographers point and North Fish Egg)
At all other seagrass meadows, 8 0.25 m2 quadrats were placed along a 100 m transect that ran perpendicular to shore at a tidal elevation between -0.5 and -0.76 m MLLW. The transect was at least 5 m (linear distance) below the upper edge of the eelgrass bed. In each quadrat, we first assessed and recorded the percent cover of macroalgae and epiphytes (two-dimensional plane within quadrat), we then counted the number of eelgrass shoots and flowering eelgrass shoots so that density could later be calculated.

In this script, we want to include only the data recorded from the quadrat measurements and create a clean ouput template to easier combine with future field season eelgrass density collections. 

Steps:
1. read in data
2. clean column names
3. combine csvs
4. combine with universal names
5. write out clean csv

```{r libraries}
library(dplyr)
library(lubridate)
library(tidyr)
library(stringr)
```

# 1. read in data
```{r data}
eg.trial <- read.csv("Data/Eelgrass_metrics/Density_RawData_2018.csv", stringsAsFactors = FALSE, 
                     header = TRUE, na.strings=c("","NA"))
eg.tran <- read.csv("Data/Eelgrass_metrics/Eelgrass_transect_2018_FINAL_WR.csv", stringsAsFactors = FALSE,
                    header = TRUE, na.strings=c("","NA"))
site_names <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Ac9c99ce9-fbdd-4879-a2c9-c90448cdba7b", method = "libcurl"), stringsAsFactors = FALSE, header = TRUE)
```


# 2. clean
```{r}
str(eg.trial)
names(eg.trial)<- tolower(names(eg.trial)) # change column names to lower case

# clean  up df and column names:
eg.trial <- eg.trial %>%
  dplyr::select(date, site, latitude, longitude, quadrat.number, blades, measurers, notes) %>%
  dplyr::rename(quadrat = quadrat.number,
                eelgrass_shoots = blades) %>% # rename column names
  separate(measurers, c("primary_observer", "secondary_observer"), sep = ",") %>% 
  # separate observer column into primary and secondary
  mutate(secondary_observer = str_trim(secondary_observer, side = "both")) %>% 
  # trim trailing/leading white spaces
  mutate(primary_observer = recode(primary_observer, 'Lia' = "LD", 'Ginny' = "GE")) %>% # use initials
  mutate(secondary_observer = recode(secondary_observer, 'Ginny' = "GE", 'Tiff' = "TS")) %>%
  mutate(macroalgae_cover = NA, # macroalgae not recorded
         diatom_cover = NA, # diatom not recorded
         flowering_shoots = NA) %>% # flowering shoots not recorded
  mutate(YYYYMMDD = mdy(date))

eg.trial$YYYYMMDD <- gsub("-", "", eg.trial$YYYYMMDD) # remove dash in date column
```

```{r}
str(eg.tran)

eg.tran <- eg.tran %>%
  dplyr::rename(macroalgae_cover = macroalgae_cover_0.25msq,
         diatom_cover = diatom_cover_0.25msq,
         eelgrass_shoots = eelgrass_shoots_0.25msq,
         flowering_shoots = flowering_shoots_0.25msq)
  

eg.tran$YYYYMMDD <- as.character(eg.tran$YYYYMMDD)
```

# 3. merge csvs
```{r}
eg.2018 <- eg.tran %>%
  bind_rows(eg.trial)
```

# 4. universal name
```{r}
eg.2018$site <- as.factor(eg.2018$site)
levels(eg.2018$site)[levels(eg.2018$site)=="Nossuk Bay-Photographer pt."] <- "Nossuk Photographers point"


eg.2018 <- eg.2018 %>%
  left_join(site_names, by = c("site" = "site_2018")) %>%
  dplyr::rename(place_name = place_name.y) %>%
  dplyr::select(site, place_name, bay_code, bay_sample, latitude.x, latitude.y, longitude.x, longitude.y, 
                -c(siteID_NOAA, site_2017, site_2019, study, freshwater, sediment_description, general_description))
```

# 5. write out csv
```{r}
write.csv(eg.2018, "Data/Eelgrass_metrics/seagrass_transect_2018_clean.csv")
```

