---
title: "Eelgrass_density+biometrics_2017"
author: "Lia Domke"
date: "5/29/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Based on scripts originally created by Wendel Raymond. 
This script takes data from two csv files:
1. Quadrat data from a 100 m transect placed through seagrass meadows. 
These data are from field season 2017 and include data from 21 sites across the western coast of Prince of Wales Island, AK. In seagrass meadows, 8, 0.25 m2 quadrats were placed along a 100 m transect that ran perpendicular to shore at a tidal elevation between -0.5 and -0.76 m MLLW. The transect was at least 5 m (linear distance) below the upper edge of the eelgrass bed. In each quadrat, we first assessed and recorded the percent cover of macroalgae and epiphytes (two-dimensional plane within quadrat), we then counted the number of eelgrass shoots and flowering eelgrass shoots so that density/m2 could later be calculated.

2. Eelgrass biometric, algal epiphytes, and epifauna. 
At the same 8 quadrats at each site, 6-7 shoots were collected carefully and placed in a mesh bag. Only 5 were used in analysis, but extras were collected as backup. For more detailed methods see metadata on KNB. 
In the lab, each bag was inverted and gently washed to remove epifauna that may have transfered to the bag. Epifauna were seperated into taxonomic groups (i.e. isopods, gammarid amphipods, caprellid amphipods, limpets, and other gastropods). We measured fresh weights for each taxonomic group and then placed the group of epifauna from each quadrat into pre-weighed foils and then were dried. The raw csv that is read in includes the freshweights, foil weights, and dried weights of epifauna and foil. 
For each of the 5 seagrass shoots (that had at least 5-cm of rhizome attached), all leaves from each of the shoots were removed and whiped clean of epiphytes using a pre-weighed and dried cotton pad (1 pad/shoot so 5 pads/quadrat and 40 pads/site). After leaves were whipped, they were individually measured for length and width. Rhizomes were standardized to 5 cm and we measured internode distance within that 5 cm segment. 
Leaves, rhizomes, epiphytes, and epifauna were all dried for at least 18 hours at 60 degrees C and weighed to the nearest 0.0001 g

Based on these two csv files the following will be calculated:
For the first file (eg.tran)

  1.1 Calculated eelgrass/flowering eelgrass shoot density/m2
  
For the second file (eg.bio)  

  2.1 dried weights for epifauna taxonomics groups will be calculated. 
  2.2 dried weights for epiphytes will be calculated
  2.3 total leaf area/shoot will be calculated (area_leaf = length*width), sum total area (leaf_areaT = sum(leaf_areas))

Then using both files can calculate quadrat level information:

  3.1 calculate aboveground biomass (shoots/m2 * shoot_mass)
      calculate belowground biomass (shoots/m2 * rhizome_mass)
      epiphytes/eelgrass area
      epiphytes per total length
      epiphytes per eelgrass mass
      grazer mass per quadrat
      grazer mass per eelgrass mass
      grazer mass/total length
      crustacean mass per eelgrass mass
      gastropod mass per eelgrass mass
      
  3.2 compile data into a single dataframe

Then calculate transect level data
  
  4.1 mean/se shoot density 
      mean/se shoot area
      mean/se shoot max
      mean/se epiphyte mass
      mean/se shoot mass
      mean/se node lengths (1-5)
      mean/se above/below ground mass
      mean/se grazer mass
      mean/se grazer mass/shoot mass
      
  4.2 merge into a single dataframe with all the data calculated at the transect level
  4.3 write out into a single csv

## Read in data, libraries, and set up standard error function
```{r libraries}
library(dplyr)
library(ggplot2)
```

```{r data}
# Site related information
eg.site <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A2c815154-9efc-44bd-b14c-be09f47f0ac1",
                        method = "libcurl"), stringsAsFactors = FALSE, header = TRUE)

site_names <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Ac9c99ce9-fbdd-4879-a2c9-c90448cdba7b", 
             method = "libcurl"), stringsAsFactors = FALSE, header = TRUE)

# Eelgrass Biometrics (lab data)
eg.bio <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Ac79aa46e-5e20-4559-86e5-64166af2cf94",
                       method = "libcurl"), stringsAsFactors = FALSE, header = TRUE)

# Eelgrass Transect (quadrat level)
eg.tran <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A5e946e41-4f5f-4499-9969-766f01113971",
                        method = "libcurl"), stringsAsFactors = FALSE, header = TRUE)
```

```{r se function}
## standard error of the mean fucntion to use later ##
st.er <- function(x, na.rm = TRUE) {
  stopifnot(is.numeric(x))
  return(sd(x, na.rm = na.rm)/sqrt(length(x)))
}
```


## Data Check
Checking the data. This is just a basic check of the data to make sure eveything looks right.
```{r checking data}
# Structure
str(eg.tran)
str(eg.site)
str(eg.bio)

# Additional code could be added here at anytime.
```

## Data prep
We want the the structure of some columns to be a certain way to that they are treated correclty for our purposes. 

Eelgrass biometrics and transect: convert plant and quadrat columns to factors
```{r number to factors}
# Transect #
eg.tran$quadrat <- as.factor(eg.tran$quadrat)

# Biometrics #
eg.bio$plant <- as.factor(eg.bio$plant)
eg.bio$quadrat <- as.factor(eg.bio$quadrat)
```

Changing start and end time column names so its not confusing
```{r changing column names}
# datachunk creator: WR

## Start and End times ##
# Site
colnames(eg.site)[6] <- "site_start_time"
colnames(eg.site)[7] <- "site_end_time"

# Transect
colnames(eg.tran)[6] <- "transect_start_time"
colnames(eg.tran)[7] <- "transect_end_time"

```

Cleaning up dates
```{r cleaning dates}
### Julian Day ###
## Biometrics ##
eg.bio$collection_date <- as.Date(eg.bio$collection_date, format = "%d/%m/%y")
eg.bio$collection_julian <- format(eg.bio$collection_date, "%j")

## Transect ##
eg.tran$date <- as.Date(eg.tran$date, format = "%m/%d/%Y")
eg.tran$date_julian <- format(eg.tran$date, "%j")

## Site ##
eg.site$date <- as.Date(eg.site$date, format = "%m/%d/%Y")
eg.site$date_julian <- format(eg.site$date, "%j")
```

## Calculations
This is broken up in chunks for clarity.
Transect, basic calculations
  1.1 Calculated eelgrass/flowering eelgrass shoot density/m2
```{r calculations - transect}
## Convert Transect data to per meter ##
eg.tran$eelgrass_shoots_msq <- (4 * eg.tran$eelgrass_shoots_0.25msq)
eg.tran$flowering_shoots_msq <- (4 * eg.tran$flowering_shoots_0.25msq)
eg.tran$macroalgae_cover_msq <- (1 * eg.tran$macroalgae_cover_0.25msq)
eg.tran$diatom_cover_msq <- (1 * eg.tran$diatom_cover_0.25msq)
```

Biometrics, basic calculations
  2.1 dried weights for epifauna taxonomics groups will be calculated. 
  2.2 dried weights for epiphytes will be calculated
  2.3 total leaf area/shoot will be calculated (area_leaf = length*width), sum total area (leaf_areaT = sum(leaf_areas))

```{r Calculations - biometrics}
### Epiphytes, Rhizomes, and Shoots Mass ###
## Epiphyte Mass ##
eg.bio$epi_mass <- (eg.bio$pad_epiphyte_mass_g - eg.bio$pad_mass_g)

## Rhizome Mass ##
eg.bio$rhi_mass <- ((eg.bio$rhi_foil_dw_g - eg.bio$rhi_foil) * (5 / eg.bio$rhi_length))

## Shoot Mass ##
eg.bio$shoot_mass <- (eg.bio$shoot_foil_dw - eg.bio$shoot_foil)

## Exess Shoot Mass ##
eg.bio$exshoot_mass <- (eg.bio$xs_shoot_foil_dw - eg.bio$xs_shoot_foil)

## Exess Epiphyte Mass ##
eg.bio$exepi_mass <- (eg.bio$xs_epiphyte_pad_mass_g - eg.bio$xs_pad_mass_g)

### Grazer Mass ###
## Isopod (Idothea rascata) Mass - bulk ##
eg.bio$iso_mass <- (eg.bio$iso_foil_dw - eg.bio$iso_foil)

## Gammarid Amphipod Mass - bulk ##
eg.bio$gamm_mass <- (eg.bio$gamm_amph_foil_dw - eg.bio$gamm_amph_foil)

## Caprellid Amphipod Mass - bulk ##
eg.bio$caprel_mass <- (eg.bio$caprel_foil_dw - eg.bio$caprel_foil)

## Limpet Mass - bulk ##
eg.bio$limp_mass <- (eg.bio$limpet_foil_dw - eg.bio$limpet_foil)

## Other Gastropod Mass - bulk ##
eg.bio$gast_mass <- (eg.bio$gastropod_foil_dw - eg.bio$gastropod_foil)

## Crab Mass ##
eg.bio$crab_mass <- (eg.bio$crab_foil_dw - eg.bio$crab_foil)

## Fill absent grazers with 0s so that cacluations below will work ##
eg.bio[,90:95][is.na(eg.bio[,90:95])] <- 0

### Effective Leaf Area ###
eg.bio$leaf_area1 <- (eg.bio$leaf_length1 * eg.bio$leaf_width1)
eg.bio$leaf_area2 <- (eg.bio$leaf_length2 * eg.bio$leaf_width2)
eg.bio$leaf_area3 <- (eg.bio$leaf_length3 * eg.bio$leaf_width3)
eg.bio$leaf_area4 <- (eg.bio$leaf_length4 * eg.bio$leaf_width4)
eg.bio$leaf_area5 <- (eg.bio$leaf_length5 * eg.bio$leaf_width5)
eg.bio$leaf_area6 <- (eg.bio$leaf_length6 * eg.bio$leaf_width6)
eg.bio$leaf_area7 <- (eg.bio$leaf_length7 * eg.bio$leaf_width7)
eg.bio$leaf_area8 <- (eg.bio$leaf_length8 * eg.bio$leaf_width8)
eg.bio$leaf_area9 <- (eg.bio$leaf_length9 * eg.bio$leaf_width9)
eg.bio$leaf_area10 <- (eg.bio$leaf_length10 * eg.bio$leaf_width10)

eg.bio$leaf_areaT <- rowSums(eg.bio[, 73:82], na.rm = TRUE)
eg.bio$leaf_max <- apply(eg.bio[, 12:21], 1, max, na.rm = TRUE)

```

Taxamonic summaries of grazers.
```{r calculations - biometrics special grazers}
## Crustacean Grazers Mass ##
eg.bio$crust_mass <- (eg.bio$iso_mass + eg.bio$gamm_mass +  eg.bio$caprel_mass + eg.bio$crab_mass)

## Gastropod Grazers Mass ##
eg.bio$gastro_mass <- (eg.bio$limp_mass + eg.bio$gast_mass)

## Total Grazer Mass ##
eg.bio$graz_massT <- rowSums(eg.bio[, 89:94], na.rm = TRUE)
```

Summarizing biometrics data by site and quadrat.
```{r biometrics - site, quadrat summaries}
## Biometrics data summarized by site and quadrat - shoots and rhizomes ##
quad.bio <- data.frame(
  eg.bio %>% 
    group_by(site, quadrat) %>% 
    summarise(s_shoot_area = mean(leaf_areaT, na.rm = TRUE),
              s_shoot_max = mean(leaf_max, na.rm = TRUE),
              s_epi_mass = mean(epi_mass, na.rm = TRUE),
              s_shoot_mass = mean(shoot_mass, na.rm = TRUE),
              s_rhi_mass = mean(rhi_mass, na.rm = TRUE),
              s_node1 = mean(node1, na.rm = TRUE),
              s_node2 = mean(node2, na.rm = TRUE),
              s_node3 = mean(node3, na.rm = TRUE),
              s_node4 = mean(node4, na.rm = TRUE),
              s_node5 = mean(node5, na.rm = TRUE),
              s_exshoot_mass = mean(exshoot_mass, na.rm = TRUE),
              s_exepi_mass = mean(exepi_mass, na.rm = TRUE))
)
```

Separating grazers. Since we pooled grazers at the quadrat level that data needs to be pulled out of the master biometrics data and summarized separatly.
```{r biometrics - grazers}
# Extract Grazer biomasses
quad.graz <- data.frame(eg.bio[eg.bio$plant == 1,]) # extract just plant 1 which has all the grazer data
quad.graz <- quad.graz[order(quad.graz[,1], quad.graz[,4]),] # order so that it matches everything else
quad.graz <- data.frame(quad.graz[,c(1, 4, 90:98)]) # reduce to get grazer data only
quad.graz[is.na(quad.graz)] <- 0 # turn NAs to 0s becasue if we sampled everything so if it was not there its a 0
```

Now we can merege transect and biometric data to summaries by site and quadrat.
```{r merge tansect and biometrics data}
## Combined Transect and Biometrics data to give site summaries ##
site.quad <- merge(eg.site[,c(1, 2)], eg.tran, by = "site")
site.quad <- merge(site.quad, quad.bio, by = c("site", "quadrat"))
site.quad <- merge(site.quad, quad.graz, by = c("site", "quadrat"))
```

Now we can preform quadrat calculations at the quadrat level.
3.1   calculate aboveground biomass (shoots/m2 * shoot_mass)
      calculate belowground biomass (shoots/m2 * rhizome_mass)
      epiphytes/eelgrass area
      epiphytes per total length
      epiphytes per eelgrass mass
      grazer mass per quadrat
      grazer mass per eelgrass mass
      grazer mass/total length
      crustacean mass per eelgrass mass
      gastropod mass per eelgrass mass
```{r calculation - quadrat level}
### Quadrat level caclulations ###
## Aboveground Biomass ##
site.quad$ag_mass <- (site.quad$eelgrass_shoots_msq * site.quad$s_shoot_mass)

## Below Ground Biomass ##
site.quad$bg_mass <- (site.quad$eelgrass_shoots_msq * site.quad$s_rhi_mass)

## Epiphytes per eeglrass area ##
site.quad$epimass_shootarea <- (site.quad$s_epi_mass / site.quad$s_shoot_area)

## Epiphytes per total length ##
site.quad$epimass_shootmax <- (site.quad$s_epi_mass / site.quad$s_shoot_max)

## Epiphytes per eelgrass mass ##
site.quad$epimass_shootmass <- (site.quad$s_epi_mass / site.quad$s_shoot_mass)

## Grazer Mass per Quadrat ##
# this is just equal to graz_massT

## Grazer mass per eelgrass mass ##
site.quad$grazmass_shootmass <- (site.quad$graz_massT / (site.quad$s_shoot_mass + site.quad$s_exshoot_mass))

## Grazer mass per total length ##
site.quad$grazmass_shootmax <- (site.quad$graz_massT / site.quad$s_shoot_max)

## Crustacean mass per eelgrass mass ##
site.quad$crustmass_shootmass <- (site.quad$crust_mass / (site.quad$s_shoot_mass + site.quad$s_exshoot_mass))

## Gastropod mass per eelgrass mass ##
site.quad$gastromass_shootmass <- (site.quad$gastro_mass / (site.quad$s_shoot_mass + site.quad$s_exshoot_mass)) 
```
## Compile data
3.2 compile data into a single dataframe
Compiling result of calculations to a usable data set.
We can now take the intermediate data frame generated above to make a complied data frame.
```{r Compiling}
#####----- Master Compile and Calculations -----#####
### Site Level Data ###
tran.dat <- eg.tran %>% 
  group_by(site) %>% 
  distinct(primary_observer, secondary_observer, transect_start_time, transect_end_time, depth_m)
  
```

Final set of calculations at site/transect level.
  4.1 mean/se shoot density 
      mean/se shoot area
      mean/se shoot max
      mean/se epiphyte mass
      mean/se shoot mass
      mean/se node lengths (1-5)
      mean/se above/below ground mass
      mean/se grazer mass
      mean/se grazer mass/shoot mass
```{r calculations - site level}
## Transect data summarized by site ##
site.tran <- data.frame(
  eg.tran %>% 
    group_by(site) %>% 
    summarise(shoot_dens = (mean(eelgrass_shoots_0.25msq * 4)),
              shoot_dens_se = (st.er(eelgrass_shoots_0.25msq * 4)),
              macro_dens = (mean(macroalgae_cover_0.25msq * 1)),
              macro_dens_se = (st.er(macroalgae_cover_0.25msq * 1)),
              diatom_dens = (mean(diatom_cover_0.25msq * 1)),
              diatom_dens_se = (st.er(diatom_cover_0.25msq * 1)),
              flower_dens = (mean(flowering_shoots_0.25msq * 4)),
              flower_dens_se = (st.er(flowering_shoots_0.25msq * 4)))
)

## Biometrics data summarized by site ##
site.bio <- data.frame(
  site.quad %>% 
    group_by(site) %>% 
    summarise(shoot_area = (mean(s_shoot_area, na.rm = TRUE)),
              shoot_area_se = (st.er(s_shoot_area)),
              shoot_max = (mean(s_shoot_max, na.rm = TRUE)),
              shoot_max_se = (st.er(s_shoot_max, na.rm = TRUE)),
              epi_mass = (mean(s_epi_mass, na.rm = TRUE)),
              epi_mass_se = (st.er(s_epi_mass)),
              shoot_mass = (mean(s_shoot_mass, na.rm = TRUE)),
              shoot_mass_se = (st.er(s_shoot_mass)),
              rhi_mass = (mean(s_rhi_mass, na.rm = TRUE)),
              rhi_mass_se = (st.er(s_rhi_mass)),
              node1 = (mean(s_node1, na.rm = TRUE)),
              node1_se = (st.er(s_node1)),
              node2 = (mean(s_node2, na.rm = TRUE)),
              node2_se = (st.er(s_node2)),
              node3 = (mean(s_node3, na.rm = TRUE)),
              node3_se = (st.er(s_node3)),
              node4 = (mean(s_node4, na.rm = TRUE)),
              node4_se = (st.er(s_node4)),
              node5 = (mean(s_node5, na.rm = TRUE)),
              node5_se = (st.er(s_node5)),
              exshoot_mass = (mean(s_exshoot_mass, na.rm = TRUE)),
              exshoot_mass_se = (st.er(s_exshoot_mass)),
              exepi_mass = (mean(s_exepi_mass, na.rm = TRUE)),
              exepi_mass_se = (st.er(s_exepi_mass)),
              abvgnd_mass = (mean(ag_mass, na.rm = TRUE)),
              abvgnd_mass_se = (st.er(ag_mass)),
              blwgnd_mass = (mean(bg_mass, na.rm = TRUE)),
              blwgnd_mass_se = (st.er(bg_mass)),
              epiphmass_shootarea = (mean(epimass_shootarea, na.rm = TRUE)),
              epiphmass_shootarea_se = (st.er(epimass_shootarea)),
              epiphmass_shootmax = (mean(epimass_shootmax, na.rm = TRUE)),
              epiphmass_shootmax_se = (st.er(epimass_shootmax, na.rm = TRUE)),
              epiphmass_shootmass = (mean(epimass_shootmass, na.rm = TRUE)),
              epiphmass_shootmass_se = (st.er(epimass_shootmass)),
              grazer_mass = (mean(graz_massT, na.rm = TRUE)),
              grazer_mass_se = (st.er(graz_massT)),
              crus_mass = (mean(crust_mass, na.rm = TRUE)),
              crus_mass_se = (st.er(crust_mass)),
              gast_mass = (mean(gastro_mass, na.rm = TRUE)),
              gast_mass_se = (st.er(gastro_mass)),
              grazermass_shootmass = (mean(grazmass_shootmass, na.rm = TRUE)),
              grazermass_shootmass_se = (st.er(grazmass_shootmass)),
              grazermass_shootmax = (mean(grazmass_shootmax, na.rm = TRUE)),
              grazermass_shootmax_se = (st.er(grazmass_shootmax)),
              crusmass_shootmass = (mean(crustmass_shootmass, na.rm = TRUE)),
              crusmass_shootmass_se = (st.er(crustmass_shootmass)),
              gastmass_shootmass = (mean(gastromass_shootmass, na.rm = TRUE)),
              gastmass_shootmass_se = (st.er(gastromass_shootmass)))
)
```

## Final merge
4.2 merge into a single dataframe with all the data calculated at the transect level
4.3 write out into a single csv
Combine everything
```{r final merge}
### Merge Transect and Biometrics data to Master dat ###
## Transect ##
dat <- site.tran %>%
  left_join(site_names, by = c("site" = "site_2017")) %>%
  dplyr::select(c(site, bay_code, bay_sample, longitude, latitude, shoot_dens, shoot_dens_se, macro_dens,
                  macro_dens_se, diatom_dens, diatom_dens_se, flower_dens, flower_dens_se))

## Biometrics ##
dat <- merge(dat, site.bio, by = "site")
```

## Export
Export data to ALL_DATA folder for use
```{r export}
#####----- Export -----#####
write.csv(dat, "Data/Eelgrass_metrics/eelgrass_and_grazer_2017_derived.csv", row.names = FALSE)
```

