---
title: "Hab_2021"
author: "Lia Domke"
date: "6/14/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

In 2021, at the same sites we conducted beach seines (see another KNB data package: https://knb.ecoinformatics.org/view/urn:uuid:04a27970-a7e7-46c8-a16b-e277a76c73c5), we also recorded data about the condition of the habitat we were seining in. We collected data from two types of habitat (eelgrass and understory kelp). 

At the eelgrass (primarily Zostera marina, but sometimes a mixture of Phyllospadix/Zostera and other times exclusively Phyllospadix) sites, we sampled eelgrass using 5 haphazardly placed quadrats (0.5 x 0.5 m) in similar area/tidal depth as where the seine net passed through. At those 5 quadrats, we counted total number of flowering and non-flowering shoots in the quadrat, and then measured the total maximum length of 15 randomly selected eelgrass plants. 

At the understory kelp sites (primarily laminaria and saacharina with *no* canopy forming kelps like nereocystsis or macrocystsis), we sampled the understory kelp using 5 haphazardly placed quadrats (0.5 x 0.5 m) in a similar area/tidal depth as where the seine net passed through. At those 5 quadrats, we estimated percent cover (%/0.25m2) of algae. In addition to percent cover, the dominant species was noted. 

At other sites still where there were adjacent understory kelp habitats we collected 5 eelgrass and 5 understory kelp quadrats for a total of 10 quadrats at a single seine site. 

After estimating percent cover, all attached algae were collected and brought back to the lab for processing. 

The algae was then identified to the lowest taxonomic level, generally species or genus level, and weighed. If the algae had a distinct holdfast and blade, we measured the total length of those individuals as well. 

These data are composed of 3 csvs files with keys to join the files together:

hab_quad -- has the quadrat level habitat data, including density, total algae biomass, and site and habitat type information

hab_lng -- has the lengths for Zostera marina/Phyllospadix and any measured algae lengths. 

hab_weight -- has the biomass and percent cover for each measured species. 

sp_list -- has the species scientific name for each year

This script is primarily for basic data checking, to make sure there aren't any errors. 

libraries
```{r}
library(tidyverse)

```

read in data - this has been visually checked and ready to be manipulated
```{r}
# 2019 data
qrt19 <- read.csv("Data/Habitat_data/hab_quadrat_2019_9-27-22.csv")
length19 <- read.csv("Data/Habitat_data/hab_length_2019_9-27-22.csv")
weight19 <- read.csv("Data/Habitat_data/hab_weight_2019_9-27-22.csv")
sp_list19 <- read.csv("Data/Habitat_data/hab_species_name_2019_9-27-22.csv")

# 2021 data, date at the end indicates download date in sept 2022 (after jess edited the 2019, 2021 hab data to make it consistent)
qrt21 <- read.csv("Data/Habitat_data/hab_quadrat_2021_9-27-22.csv")
length21 <- read.csv("Data/Habitat_data/hab_length_2021_9-27-22.csv")
weight21 <- read.csv("Data/Habitat_data/hab_weight_2021_9-27-22.csv")
sp_list21 <- read.csv("Data/Habitat_data/hab_species_name_2021_9-27-22.csv")

# 2022 data, downloaded in march 2023 (jess entered and edited to make consistent with 2019, 2021 - in 08-2022)
# updated some mistaken quadrat numbers (3-14-23)
qrt22 <- read.csv("Data/Habitat_data/hab_quadrat_2022_03-14-23.csv")
length22 <- read.csv("Data/Habitat_data/hab_length_2022_03-14-23.csv")
weight22 <- read.csv("Data/Habitat_data/hab_weight_2022_03-14-23.csv")
sp_list22 <- read.csv("Data/Habitat_data/hab_species_name_2022_03-14-23.csv")

# site.ids and names
sites <- read.csv("Data/Site_metadata_allyears_8-18-22.csv")

```

Want to check to make sure there aren't any errors, but first lets join the columns together. 
That means we need to: 

- add in bay_id to the 2019 dataset
- match the column names 
- add in any units missing in column titles
- add in weight_g column to 2019 (even tho there won't be any measurements)
- in each year, make sure the species listed are present (and correctly spelled) in the species is column. 
- calculate total_biomass for each quadrat and make sure its correct
- when joining together the datasets, each df has the key (in this case quadrat_ID), in each year we started over at 1. In order to combine across years we need to either combine the quadratID with the year or re do quadrat numbers

```{r}
glimpse(qrt19)
unique(qrt19[,c("quadrat_ID", "site", "date", "YYYYMMDD", "notes")]) # looks pretty good
unique(qrt19$quadrat_ID) # 125 quadrats representing 25 sites

glimpse(qrt21)
unique(qrt21[,c("quadrat_ID", "site", "bay_ID", "date", "YYYYMMDD")]) # looks pretty good
unique(qrt21$quadrat_ID) # 115 quadrats representing 23 sites (not all were resampling of noaa sites)

glimpse(qrt22)
unique(qrt22[,c("quadrat_ID", "site", "bay_ID", "date", "YYYYMMDD")])
unique(qrt22$quadrat_ID) # 105 representing 13 sites
```

add in bay id to 2019 dataset
```{r}
unique(qrt19$site)

qrt19$site <- as.factor(qrt19$site)
levels(qrt19$site)[levels(qrt19$site)=="Goats mouth inlet"]<-"Goat Mouth Inlet"
levels(qrt19$site)[levels(qrt19$site)=="Naukati Bay"]<-"Naukati"
levels(qrt19$site)[levels(qrt19$site)=="South Wadleigh Island"]<-"South Wadleigh"
levels(qrt19$site)[levels(qrt19$site)=="Guktu Bay"]<-"Guktu"
levels(qrt19$site)[levels(qrt19$site)=="North Fish Egg Island"]<-"North Fish Egg"
levels(qrt19$site)[levels(qrt19$site)=="North Fish Egg - kelp"]<-"North Fish Egg-Kelp"
levels(qrt19$site)[levels(qrt19$site)=="Natzuhini Bay"]<-"Natzuhini"
levels(qrt19$site)[levels(qrt19$site)=="Kaguk Cove"]<-"Kaguk "
levels(qrt19$site)[levels(qrt19$site)=="Farallon Bay"]<-"Farallon"
levels(qrt19$site)[levels(qrt19$site)=="Chusini Cove"]<-"Chusini-Kladein Flat"
levels(qrt19$site)[levels(qrt19$site)=="South Fish Egg Island"]<-"South Fish Egg"
levels(qrt19$site)[levels(qrt19$site)=="Baker Island - kelp"]<-"Baker Island - kelp "
levels(qrt19$site)[levels(qrt19$site)=="Shinaku Inlet "] <- "Shinaku Inlet"
qrt19$site <- as.character(qrt19$site)

qrt19.2 <- qrt19 %>%
  left_join(sites, by = c("site" = "site_2019")) %>%
  unite(bay_id, bay_code:bay_sample) %>%
  mutate(habitat_quadrat = habitat.y,
         habitat_site = habitat.x) %>%
  dplyr::select(-c(freshwater, sediment_description, general_description, habitat.x, habitat.y, site_2017, site_2018, study))
  

unique(qrt19.2[,c("site", "date", "YYYYMMDD", "bay_id", "habitat_quadrat", "siteID_NOAA")]) # looks pretty good

```

do the quadrat column names match between 21 and 19
```{r}
names(qrt19.2)
names(qrt21)
site.id <- unite(sites, bay_id, bay_code:bay_sample)

qrt19.3 <- qrt19.2 %>%
  dplyr::rename(total_biomass_g = total_biomass, density_0.25msq = density,
                flowering_0.25msq = flowering_shoots) %>%
  mutate(depth_ft = NA, time = NA)

qrt21.2 <- qrt21 %>%
  left_join(site.id, by = c("bay_ID" = "bay_id")) %>%
  dplyr::rename(habitat_quadrat = habitat.x, habitat_site = habitat.y, bay_id = bay_ID) %>%
  dplyr::select(-c(X, X.1, site_2017, site_2018, site_2019, freshwater, sediment_description, general_description, study)) %>%
  mutate(depth_ft = NA, time = NA)

names(qrt21.2)

unique(qrt21.2[,c("site", "date", "YYYYMMDD", "bay_id", "habitat_quadrat", "habitat_site", "siteID_NOAA")]) # looks pretty good

# adjust columns for 2022
# which sites are mixed?
sites.hab <- sites %>%
  unite(bay_id, bay_code:bay_sample) %>%
  dplyr::select(bay_id, habitat, siteID_NOAA, latitude, longitude, place_name)

qrt22.1 <- qrt22 %>%
  dplyr::rename(bay_id = bay_ID,
         habitat_quadrat = habitat) %>%
  left_join(sites.hab, by = "bay_id") %>%
  dplyr::rename(habitat_site = habitat)
  


names(qrt19.3)
names(qrt21.2)
names(qrt22.1) # check, names match!


```

- add in weight_g column to 2019 (even tho there won't be any measurements)
```{r}
glimpse(length19)
glimpse(length21)
glimpse(length22)

length19.2 <- length19 %>%
  dplyr::rename(length_mm = length) %>%
  mutate(weight_g = "NA")

# both match column names!
names(length19.2); names(length21); names(length22)
```

- in each year, make sure the species listed are present (and correctly spelled) in the species is column. 
```{r}
sp_list19$species.scientific.name <- trimws(sp_list19$species.scientific.name)

sp_list19 %>%
  arrange()

# raname the species that aren't correct

length19.3 <- length19.2 %>%
  mutate(species = ifelse(species == "Cumantamnion decipiens", "Cumanthamnion decipiens", species),
         species = ifelse(species == "Desmersita spp", "Desmarestia spp", species),
         species = ifelse(species == "Farlomia mollis" | species == "Falonia mollis", "Farlowia mollis", species),
         species = ifelse(species == "Phyllospadix" | species == "Phyllospadix scouleri", "Phyllospadix serratus", species), # this one is because likely this was a mistake in identification. Phyllospadix scouleri does not occur mixed with eelgrass, Phyllospadix serratus does. 
         species = ifelse(species == "Saccharina groelandica", "Saccharina groenlandica", species),
         species = ifelse(species == "Sargassum", "Sargassum spp", species),
         species = ifelse(species == "Sparangia pertula", "Sparlingia pertusa", species),
         species = ifelse(species == "Ulva lactuca ", "Ulva lactuca", species),
         species = ifelse(species == "Ulva Spp", "Ulva spp", species),
         species = ifelse(species == "Saacharina spp", "Saccharina spp", species))

weight19.2 <- weight19 %>%
   mutate(species = ifelse(species == "Cumantamnion decipiens", "Cumanthamnion decipiens", species),
         species = ifelse(species == "Desmersita spp", "Desmarestia spp", species),
         species = ifelse(species == "Farlomia mollis" | species == "Falonia mollis", "Farlowia mollis", species),
         species = ifelse(species == "Phyllospadix" | species == "Phyllospadix scouleri", "Phyllospadix serratus", species),
         species = ifelse(species == "Saccharina groelandica", "Saccharina groenlandica", species),
         species = ifelse(species == "Sargassum", "Sargassum spp", species),
         species = ifelse(species == "Sparangia pertula", "Sparlingia pertusa", species),
         species = ifelse(species == "Ulva lactuca ", "Ulva lactuca", species),
         species = ifelse(species == "Ulva Spp", "Ulva spp", species),
         species = ifelse(species == "Saacharina spp", "Saccharina spp", species)) %>%
  rename(percent_cover = X._cover,
         weight_g = weight)

species19 <- length19.3 %>%
  dplyr::select(species) %>%
  distinct()

species19 <- weight19.2 %>%
  dplyr::select(species) %>%
  distinct() %>%
  rbind(species19) %>%
  distinct() %>%
  arrange(species)

anti_join(sp_list19, species19, by = c("species.scientific.name" = "species")) # only one species on sp_list that doesn't occur in the dataset, its the phyllospadix scouleri that we removed above. 

# create new species list that includes all the species that occur in the data set
# remove the sediment and non species (like biomass)
remove <- c("bare", "bare rock", "bare rock/sediment exposed/uncovered", "Biomass", "course sand w/green algae", "Micellaneous", "rock",
            "shell hash/boulders")
list19.2 <- full_join(species19, sp_list19, by = c("species" = "species.scientific.name")) %>%
  subset(!(species %in% remove)) %>%
  filter(species != "Phyllospadix scouleri")
```

Check names in 2021 too
```{r}
sp_list21$species_scientific_name <- trimws(sp_list21$species_scientific_name)
sp_list22$species_scientific_name <- trimws(sp_list22$species_scientific_name)

sp_list21 <- sp_list21 %>%
  dplyr::select(-alternative_species_names) %>%
  arrange()

sp_list22 <- sp_list22 %>%
  dplyr::select(-alternative_species_names) %>%
  arrange()

species21 <- length21 %>%
  dplyr::select(species) %>%
  distinct()

species22 <- length22 %>%
  dplyr::select(species) %>%
  distinct()

species21 <- weight21 %>%
  dplyr::select(species) %>%
  distinct() %>%
  rbind(species21) %>%
  distinct() %>%
  arrange(species)

species22 <- weight22 %>%
  dplyr::select(species) %>%
  distinct() %>%
  rbind(species22) %>%
  distinct() %>%
  arrange(species)

anti_join(sp_list21, species21, by = c("species_scientific_name" = "species")) # a few species missing 
missing21 <- species21 %>%
  dplyr::select(species) %>%
  arrange(species) %>%
  full_join(sp_list21, by = c("species" = "species_scientific_name"), keep = TRUE, na_matches = "na")

anti_join(sp_list22, species22, by = c("species_scientific_name" = "species")) # a few species missing 

missing22 <- species22 %>%
  dplyr::select(species) %>%
  arrange(species) %>%
  full_join(sp_list22, by = c("species" = "species_scientific_name"), keep = TRUE, na_matches = "na")

# change the names to make sure that there is consistency between data
length21$species <- trimws(length21$species)
length21.2 <- length21 %>%
  mutate(species = ifelse(species == "Acrosiphinia arcta", "Acrosiphonia arcta", species),
         species = ifelse(species == "bare/shell", "bare/shells", species),
         species = ifelse(species == "branch red", "branching red", species),
         species = ifelse(species == "cobble", "Cobble", species), 
         species = ifelse(species == "eelgrass" | species == "Zostera marina " | species == "seagrass", "Zostera marina", species),
         species = ifelse(species == "eudesme virescens", "Eudesme virescens", species), 
         species = ifelse(species == "Fucus gardneri", "Fucus distichus", species),
         species = ifelse(species == "macrocystis pyrifera", "Macrocystis pyrifera", species), 
         species = ifelse(species == "saccharina spp", "Saccharina spp", species),
         species = ifelse(species == "sand/shells", "Sand/shells", species),
         species = ifelse(species == "cobble/sand", "Cobble/sand", species),
         species = ifelse(species == "sand", "Sand", species),
         species = ifelse(species == "Shells/sand", "Sand/shells", species),
         species = ifelse(species == "Schizymena pacifica", "Schizymenia pacifica", species))

weight21$species <- trimws(weight21$species)

weight21.2 <- weight21 %>%
  mutate(species = ifelse(species == "Acrosiphinia arcta", "Acrosiphonia arcta", species),
         species = ifelse(species == "bare/shell", "bare/shells", species),
         species = ifelse(species == "branch red", "branching red", species),
         species = ifelse(species == "cobble", "Cobble", species), 
         species = ifelse(species == "eelgrass" | species == "seagrass", "Zostera marina", species),
         species = ifelse(species == "eudesme virescens", "Eudesme virescens", species), 
         species = ifelse(species == "Fucus gardneri", "Fucus distichus", species),
         species = ifelse(species == "macrocystis pyrifera", "Macrocystis pyrifera", species), 
         species = ifelse(species == "saccharina spp", "Saccharina spp", species),
         species = ifelse(species == "sand/shells", "Sand/shells", species),
         species = ifelse(species == "cobble/sand", "Cobble/sand", species),
         species = ifelse(species == "sand", "Sand", species),
         species = ifelse(species == "Shells/sand", "Sand/shells", species),
         species = ifelse(species == "Schizymena pacifica", "Schizymenia pacifica", species))

remove21 <- c("bare", "bare and shell hash", "bare cobble/pebble", "bare/dead kelp", "bare/shells", 
            "bare rock/sediment exposed/uncovered", "boulder", "Cobble", "Cobble/boulder",
            "Cobble/pebbles", "Cobble/sand", "gravel", "large rock", "mud", "Sand", "Sand/shells")


length22$species <- trimws(length22$species)
length22.2 <- length22 %>%
  mutate(species = ifelse(species == "Acrosiphinia arcta" | species == "Arcosiphonia arcta", "Acrosiphonia arcta", species),
         species = ifelse(species == "Cymathaene triplicata", "Cymathaere triplicata",species),
         species = ifelse(species == "Desmarestia acuelata", "Desmarestia aculeata", species),
         species = ifelse(species == "Mazzaella phylocarpa", "Mazzaella phyllocarpa", species),
         species = ifelse(species == "Neorhodomela acuelata", "Neorhodomela aculeata", species),
         species = ifelse(species == "cobble", "Cobble", species),
         species = ifelse(species == "boulder", "Boulder", species),
         species = ifelse(species == "sand", "Sand", species),
         species = ifelse(species == "corallina officinalis", "Corallina officinalis", species),
         species = ifelse(species == "dictyosiphon foeniculaceus", "Dictyosiphon foeniculaceus", species),
         species = ifelse(species == "farlowia mollis", "Farlowia mollis", species),
         species = ifelse(species == "fucus gardenari", "Fucus gardenari", species),
         species = ifelse(species == "leathesia marina", "Leathesia marina", species),
         species = ifelse(species == "mazzaella sp", "Mazzaella sp", species),
         species = ifelse(species == "mazzaella splendens", "mazzaella splendens", species),
         species = ifelse(species == "neodilsea borealis", "Neodilsea borealis", species),
         species = ifelse(species == "sacharina groenlandica", "Sacharina groenlandica", species),
         species = ifelse(species == "saccharina spp", "Saccharina spp", species),
         species = ifelse(species == "saccharina spp biomass", "Saccharina spp biomass", species),
         species = ifelse(species == "sarcodiotheca gaudichaudii", "Sarcodiotheca gaudichaudii", species),
         species = ifelse(species == "sphacelaria spp", "Sphacelaria spp", species),
         species = ifelse(species == "ulva lactuca", "Ulva lactuca", species),
         species = ifelse(species == "ulvaria sp" | species == "ulvaria spp", "Ulvaria spp", species))

weight22$species <- trimws(weight22$species)
weight22.2 <- weight22 %>%
  mutate(species = ifelse(species == "Acrosiphinia arcta" | species == "Arcosiphonia arcta", "Acrosiphonia arcta", species),
         species = ifelse(species == "Cymathaene triplicata", "Cymathaere triplicata",species),
         species = ifelse(species == "Desmarestia acuelata", "Desmarestia aculeata", species),
         species = ifelse(species == "Mazzaella phylocarpa", "Mazzaella phyllocarpa", species),
         species = ifelse(species == "Neorhodomela acuelata", "Neorhodomela aculeata", species),
         species = ifelse(species == "cobble", "Cobble", species),
         species = ifelse(species == "boulder", "Boulder", species),
         species = ifelse(species == "sand", "Sand", species),
         species = ifelse(species == "corallina officinalis", "Corallina officinalis", species),
         species = ifelse(species == "dictyosiphon foeniculaceus", "Dictyosiphon foeniculaceus", species),
         species = ifelse(species == "farlowia mollis", "Farlowia mollis", species),
         species = ifelse(species == "fucus gardenari", "Fucus gardenari", species),
         species = ifelse(species == "leathesia marina", "Leathesia marina", species),
         species = ifelse(species == "mazzaella sp", "Mazzaella sp", species),
         species = ifelse(species == "mazzaella splendens", "mazzaella splendens", species),
         species = ifelse(species == "neodilsea borealis", "Neodilsea borealis", species),
         species = ifelse(species == "sacharina groenlandica", "Sacharina groenlandica", species),
         species = ifelse(species == "saccharina spp", "Saccharina spp", species),
         species = ifelse(species == "saccharina spp biomass", "Saccharina spp biomass", species),
         species = ifelse(species == "sarcodiotheca gaudichaudii", "Sarcodiotheca gaudichaudii", species),
         species = ifelse(species == "sphacelaria spp", "Sphacelaria spp", species),
         species = ifelse(species == "ulva lactuca", "Ulva lactuca", species),
         species = ifelse(species == "ulvaria sp" | species == "ulvaria spp", "Ulvaria spp", species))

remove22 <- c("bare", "bare and shell hash", "bare cobble/pebble", "bare/dead kelp", "bare/shells", 
            "bare rock/sediment exposed/uncovered", "boulder", "Cobble", "Cobble/boulder",
            "Cobble/pebbles", "Cobble/sand", "gravel", "large rock", "mud", "Sand", "Sand/shells")

#_____________________
species21.2 <- length21.2 %>%
  dplyr::select(species) %>%
  distinct()

species21.3 <- weight21.2 %>%
  dplyr::select(species) %>%
  distinct() %>%
  rbind(species21.2) %>%
  distinct() %>%
  arrange(species)

list21.2 <- species21.3 %>%
  subset(!(species %in% remove21))
anti_join(list21.2, species21.3, by = "species") # no species!

#--------------

species22.2 <- length22.2 %>%
  dplyr::select(species) %>%
  distinct()

species22.3 <- weight22.2 %>%
  dplyr::select(species) %>%
  distinct() %>%
  rbind(species22.2) %>%
  distinct() %>%
  arrange(species)

list22.2 <- species22.3 %>%
  subset(!(species %in% remove22))
anti_join(list21.2, species21.3, by = "species") # no species!
```


- calculate total_biomass for each quadrat and make sure its correct: 2019
```{r}
head(qrt19.3)
head(length19.3)
glimpse(weight19.2)

# no weights in the length (that would've been invidual plant weights)
total.biom19 <- weight19.2 %>%
  group_by(quadrat_ID) %>%
  dplyr::summarise(total_biomass_g2 = sum(weight, na.rm = TRUE))

g_check <- qrt19.3 %>%
  filter(habitat_quadrat == "kelp") %>%
  left_join(total.biom19, by = "quadrat_ID") 

g_check$check <- g_check$total_biomass_g == g_check$total_biomass_g2
g_check$g_diff <- round(g_check$total_biomass_g - g_check$total_biomass_g2, digits = 2)

filter(g_check, check == FALSE) # all these cases are either rounding errors or a legit calculation error. Go with the R calc ##

qrt19.4 <- qrt19.3 %>%
  left_join(total.biom19, by = "quadrat_ID") %>%
  dplyr::select(-total_biomass_g) %>%
  rename(total_biomass_g = total_biomass_g2)
```

- calculate total_biomass for each quadrat and make sure its correct: 2021
```{r}
head(qrt21.2)
head(length21.2)
glimpse(weight21.2)

# 2021 has weights associated with the individual lengths of certain plants (mainly phyllospadix). And this is eventually to develop a length to weight relationship if need to better est biomass of phyllospadix meadows or mixed phyllo/zos meadows
weight21.2$weight_g <- as.numeric(weight21.2$weight_g)

total.biom21 <- weight21.2 %>%
  group_by(quadrat_ID) %>%
  dplyr::summarise(total_biomass_g_2 = sum(weight_g, na.rm = TRUE))

g_check <- qrt21.2 %>%
  left_join(total.biom21, by = "quadrat_ID") 

g_check$total_biomass_g <- as.numeric(g_check$total_biomass_g)
g_check$check <- g_check$total_biomass_g == g_check$total_biomass_g_2
g_check$g_diff <- round(g_check$total_biomass_g - g_check$total_biomass_g_2, digits = 2)

filter(g_check, check == FALSE) # all these cases are either rounding errors or a legit calculation error. Go with the R calc ##

qrt21.3 <- qrt21.2 %>%
  left_join(total.biom21, by = "quadrat_ID") %>%
  dplyr::select(-total_biomass_g) %>%
  rename(total_biomass_g = total_biomass_g_2)


#### because there are lengths associated with weights lets make sure they're correct
weight.sp.quad21 <- length21.2 %>%
  filter(!(is.na(weight_g))) %>%
  group_by(quadrat_ID, species) %>%
  dplyr::summarise(weight_g = sum(weight_g))

weight.sp.quad21$X._cover <- NA
weight.sp.quad21$species_by_quadrat_ID <- seq(from = 471, length.out = 9)
weight.sp.quad21$notes <- NA

#lets add in an entry that didn't make it from clam island
missingwgt <- setNames(data.frame(matrix(ncol = 6, nrow = 0)), c("quadrat_ID", "species", "weight_g", "percent_cover", "species_by_quadrat_ID", "notes"))
missingwgt <- rbind(missingwgt, c("94", "Zostera marina", "NA", "100", "476", "NA"))
colnames(missingwgt) <- c("quadrat_ID", "species", "weight_g", "percent_cover", "species_by_quadrat_ID", "notes")
missingwgt$quadrat_ID <- as.integer(missingwgt$quadrat_ID)
missingwgt$weight_g <- as.numeric(missingwgt$weight_g)
missingwgt$species_by_quadrat_ID <- as.numeric(missingwgt$species_by_quadrat_ID)

weight21.3.1 <- weight.sp.quad21 %>%
  filter(quadrat_ID < 76) %>%
  rbind(weight21.2) %>%
  rename(percent_cover = X._cover)

weight21.3.2 <- rbind(weight21.3.1, missingwgt)

# lets adjust an instance where MOIR_B had the wrong percent cover numbers
weight21.3 <- mutate(weight21.3.2, percent_cover = ifelse(quadrat_ID == 61 & species == "boulder" & percent_cover == 33, 23, percent_cover))


## make sure it all matches up
quad.weights <- weight21.3 %>%
  subset(quadrat_ID %in% weight.sp.quad21$quadrat_ID) %>%
  group_by(quadrat_ID) %>%
  dplyr::summarise(sum = sum(weight_g)) %>%
  filter(quadrat_ID < 76)

qrt21.4 <- qrt21.3 %>%
  left_join(quad.weights)
```

- calculate total_biomass for each quadrat and make sure its correct: 2022
```{r}
head(qrt22.1)
head(length22.2)
glimpse(weight22.2)

# 2021 has weights associated with the individual lengths of certain plants (mainly phyllospadix). And this is eventually to develop a length to weight relationship if need to better est biomass of phyllospadix meadows or mixed phyllo/zos meadows


# HI LIA THERE ARE DEFINITELY SOME INSTANCES WHERE THE DIFFERENCES IN WEIGHT ARE SUBSTANTIAL. 
# Blanquizal quadrat ID 60, quadrat 5 in kelp has a .2 g diff - macrocystis biomass should be 28.6 NOT 28.8
# Halibut nose quadrat 94, quadrat 4 has a 5 gram diff - it says that cobble is 5 g, but its not 
# the rest are looking like rounding errors. 

qrt22.2 <- qrt22.1 

weight22.3 <- weight22.2 %>%
  mutate(weight_g = ifelse(species_by_quadrat_ID == 260 & species == "Macrocystis pyrifera biomass", 28.6, weight_g),
         weight_g = ifelse(species_by_quadrat_ID == 313 & species == "Cobble", NA, weight_g)) %>%
  # theres a duplicate entry of neodilsea borealis 50.3 that needs to be removed from SARI_A
  filter(!(quadrat_ID == 36 & species == "Neodilsea borealis" & X._cover == "not collected" & species_by_quadrat_ID == 147)) # one less row good. 

weight22.3$weight_g <- as.numeric(weight22.3$weight_g)

total.biom22 <- weight22.3 %>%
  group_by(quadrat_ID) %>%
  dplyr::summarise(total_biomass_g_2 = sum(weight_g, na.rm = TRUE))

g_check <- qrt22.2 %>%
  left_join(total.biom22, by = "quadrat_ID") 

g_check$total_biomass_g <- as.numeric(g_check$total_biomass_g)
g_check$check <- g_check$total_biomass_g == g_check$total_biomass_g_2
g_check$g_diff <- round(g_check$total_biomass_g - g_check$total_biomass_g_2, digits = 2)

filter(g_check, check == FALSE) # all rounding errors
##############
qrt21.3 <- qrt21.2 %>%
  left_join(total.biom21, by = "quadrat_ID") %>%
  dplyr::select(-total_biomass_g) %>%
  rename(total_biomass_g = total_biomass_g_2)


#### because there are lengths associated with weights lets make sure they're correct
weight.sp.quad21 <- length21.2 %>%
  filter(!(is.na(weight_g))) %>%
  group_by(quadrat_ID, species) %>%
  dplyr::summarise(weight_g = sum(weight_g))

weight.sp.quad21$X._cover <- NA
weight.sp.quad21$species_by_quadrat_ID <- seq(from = 471, length.out = 9)
weight.sp.quad21$notes <- NA


#lets add in an entry that didn't make it from clam island
missingwgt <- setNames(data.frame(matrix(ncol = 6, nrow = 0)), c("quadrat_ID", "species", "weight_g", "percent_cover", "species_by_quadrat_ID", "notes"))
missingwgt <- rbind(missingwgt, c("94", "Zostera marina", "NA", "100", "476", "NA"))
colnames(missingwgt) <- c("quadrat_ID", "species", "weight_g", "percent_cover", "species_by_quadrat_ID", "notes")
missingwgt$quadrat_ID <- as.integer(missingwgt$quadrat_ID)
missingwgt$weight_g <- as.numeric(missingwgt$weight_g)
missingwgt$species_by_quadrat_ID <- as.numeric(missingwgt$species_by_quadrat_ID)

weight21.3.1 <- weight.sp.quad21 %>%
  filter(quadrat_ID < 76) %>%
  rbind(weight21.2) %>%
  rename(percent_cover = X._cover)

weight21.3.2 <- rbind(weight21.3.1, missingwgt)

# lets adjust an instance where MOIR_B had the wrong percent cover numbers
weight21.3 <- mutate(weight21.3.2, percent_cover = ifelse(quadrat_ID == 61 & species == "boulder" & percent_cover == 33, 23, percent_cover))


## make sure it all matches up
quad.weights <- weight21.3 %>%
  subset(quadrat_ID %in% weight.sp.quad21$quadrat_ID) %>%
  group_by(quadrat_ID) %>%
  dplyr::summarise(sum = sum(weight_g)) %>%
  filter(quadrat_ID < 76)

qrt21.4 <- qrt21.3 %>%
  left_join(quad.weights)
```

remove any of the df that we're not interested in
```{r}
rm(list = c("g_check", "length19", "length19.2","qrt19", "qrt19.2", 
            "qrt19.3", "qrt21", 
            "sp_list19", "species19", 
            "weight19", 
            "length21", "weight21", "sp_list21", "species21", "species21.2", "species21.3", 
            "weight21.2",
            "length22", "qrt22", "qrt22.1", "weight22.2"))
```

- when joining together the datasets, each df has the key (in this case quadrat_ID), in each year we started over at 1. In order to combine across years we need to either combine the quadratID with the year or re do quadrat numbers

```{r}
# the data from 2019 includes
head(qrt19.4)
head(length19.3)
head(weight19.2)
head(list19.2)

head(qrt21.3)
head(length21.2)
head(weight21.3)
head(list21.2)

head(qrt22.2)
head(length22.2)
head(weight22.3)
head(list22.2)

names(qrt19.4)
names(qrt21.3)
names(qrt22.2)

names(length19.3)
names(length21.2)
names(length22.2)

names(weight19.2)
names(weight21.3)
names(weight22.3)
# for now (10/11/22) lets read out the cleanish data from 2021 
#write.csv(qrt21.3, "Data/Habitat_data/hab_quadrat_2021_clean_10-11-22.csv")
#write.csv(length21.2, "Data/Habitat_data/hab_length_2021_clean_10-11-22.csv")
#write.csv(weight21.3, "Data/Habitat_data/hab_weight_2021_clean_10-11-22.csv")
#write.csv(list21.2, "Data/Habitat_data/hab_sp_list_2021_clean_10-11-22.csv")

#write.csv(qrt19.4, "Data/Habitat_data/hab_quadrat_2019_clean_10-11-22.csv")
#write.csv(length19.3, "Data/Habitat_data/hab_length_2019_clean_10-11-22.csv")
#write.csv(weight19.2, "Data/Habitat_data/hab_weight_2019_clean_10-11-22.csv")
#write.csv(list19.2, "Data/Habitat_data/hab_sp_list_2019_clean_10-11-22.csv")

#write.csv(qrt22.2, "Data/Habitat_data/hab_quadrat_2022_clean_3-14-23.csv")
#write.csv(length22.2, "Data/Habitat_data/hab_length_2022_clean_3-14-23.csv")
#write.csv(weight22.3, "Data/Habitat_data/hab_weight_2022_clean_3-14-23.csv")
#write.csv(list22.2, "Data/Habitat_data/hab_sp_list_2022_clean_3-14-23.csv")


```

