---
title: "Hab_2019"
author: "Lia Domke"
date: "6/10/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In 2019, at the same sites we conducted beach seines (see another KNB data package: https://knb.ecoinformatics.org/view/urn:uuid:04a27970-a7e7-46c8-a16b-e277a76c73c5), we also recorded data about the condition of the habitat we were seining in. We collected data from two types of habitat (eelgrass and understory kelp). 

At the eelgrass (Zostera marina) sites, we sampled eelgrass using 5 haphazardly placed quadrats (0.5 x 0.5 m) in similar area/tidal depth as where the seine net passed through. At those 5 quadrats, we counted total number of flowering and non-flowering shoots in the quadrat, and then measured the total maximum length of 15 randomly selected eelgrass plants. 

At the understory kelp sites (primarily laminaria and saacharina with *no* canopy forming kelps like nereocystsis or macrocystsis), we sampled the understory kelp using 5 haphazardly placed quadrats (0.5 x 0.5 m) in a similar area/tidal depth as where the seine net passed through. At those 5 quadrats, we estimated percent cover (%/0.25m2) of algae. In addition to percent cover, the dominant species was noted. 
After estimating percent cover, all attached algae were collected and brought back to the lab for processing. 

The algae was then identified to the lowest taxonomic level, generally species or genus level, and weighed. If the algae had a distinct holdfast and blade, we measured the total length of those individuals as well. 

These data are composed of 3 xlsx files with keys to join the files together:
hab_quad -- has the quadrat level habitat data, including density, total algae biomass, and site and habitat type information

hab_lng -- has the lengths for Zostera marina and any measured algae lengths. 

hab_biom -- has the biomass and percent cover for each measured species. 

This script is primarily for basic data checking, to make sure there aren't any errors. 

```{r libraries}
library(dplyr)
library(readxl)
```

```{r data}
hab_quad <- read_xlsx("Data/Habitat_data/Habitat_2019_RAW_2-21-2020.xlsx", sheet = 1)
hab_lng <- read_xlsx("Data/Habitat_data/Habitat_2019_RAW_2-21-2020.xlsx", sheet = 2)
hab_biom <- read_xlsx("Data/Habitat_data/Habitat_2019_RAW_2-21-2020.xlsx", sheet = 3)

# universal site information
loc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Ac9c99ce9-fbdd-4879-a2c9-c90448cdba7b"), stringsAsFactors = FALSE, header = TRUE)
```

Functions
```{r standar error funciton}
## standard error of the mean fucntion to use later ##
st.er <- function(x, na.rm = TRUE) {
  stopifnot(is.numeric(x))
  return(sd(x, na.rm = na.rm)/sqrt(length(x)))
}
```

Mainly need to do some basic data checking to make sure that the data connect properly and don't have any errors. 
```{r}
str(hab_quad)
str(hab_lng)
str(hab_biom)
```


```{r}
unique(hab_quad$site) # 25 sites, correct
unique(hab_lng$quadrat_ID) # 125 quadrats, correct
length(unique(hab_biom$quadrat_ID)) # quadrats from 25 sites, correct
unique(hab_quad$date)
```

Change seagrass to eelgrass
```{r}
hab_quad$habitat <- as.factor(hab_quad$habitat)
levels(hab_quad$habitat)[levels(hab_quad$habitat)=="seagrass"] <- "eelgrass"
hab_quad$habitat <- as.character(hab_quad$habitat)
```

Rounding issue
```{r}
str(hab_quad)
hab_quad$total_biomass <- as.numeric(hab_quad$total_biomass)
round(hab_quad$total_biomass, 1)

str(hab_biom) # change the rest of the columns from character to numeric
hab_biom$'%_cover' <- as.numeric(hab_biom$'%_cover')
hab_quad$density <- as.numeric(hab_quad$density)
hab_quad$flowering_shoots <- as.numeric(hab_quad$flowering_shoots)
```

Summarize data to site level
```{r}
site_level <- hab_quad %>%
  group_by(site) %>%
  mutate(avg.biom = mean(total_biomass)) %>%
  mutate(biom.se = st.er(total_biomass)) %>%
  mutate(avg.density = mean(density)) %>%
  mutate(density.se = st.er(density)) %>%
  mutate(avg.flowering = mean(flowering_shoots)) %>%
  mutate(flowering.se = mean(flowering_shoots)) %>%  
  dplyr::select(-c(density, flowering_shoots, total_biomass, quadrat, quadrat_ID, notes)) %>%
  distinct()
```

Calculate quadrat level
- calculate average length per species per quadrat
```{r}

quad_level <- hab_lng %>%
  group_by(quadrat_ID, species) %>% # average/se by species and quadrat
  mutate(avg.length  = mean(length)) %>%
  mutate(length.se = st.er(length)) %>%
  dplyr::select(-c(length, plant_ID)) %>%
  distinct() %>%
  left_join(hab_biom) %>% # left join with biomass/% cover measurements
  dplyr::select(-c(species_by_quadrat_ID)) %>%
  full_join(hab_quad) %>% # join with other quadrat data (total biomass)
  dplyr::select(c(site, date, YYYYMMDD, habitat, quadrat_ID, species, avg.length, length.se,
                  weight, '%_cover', total_biomass, density, flowering_shoots)) # change order of columns
  
```

Adjust quadrat data sheet to include bay id and bay code for ease of combination
```{r}
# the names are slightly off to be able to combine with the location df
hab_quad$site <- as.factor(hab_quad$site)
levels(hab_quad$site)[levels(hab_quad$site)=="Goats mouth inlet"]<-"Goat Mouth Inlet"
levels(hab_quad$site)[levels(hab_quad$site)=="Naukati Bay"]<-"Naukati"
levels(hab_quad$site)[levels(hab_quad$site)=="South Wadleigh Island"]<-"South Wadleigh"
levels(hab_quad$site)[levels(hab_quad$site)=="Guktu Bay"]<-"Guktu"
levels(hab_quad$site)[levels(hab_quad$site)=="North Fish Egg Island"]<-"North Fish Egg"
levels(hab_quad$site)[levels(hab_quad$site)=="North Fish Egg - kelp"]<-"North Fish Egg-Kelp"
levels(hab_quad$site)[levels(hab_quad$site)=="Natzuhini Bay"]<-"Natzuhini"
levels(hab_quad$site)[levels(hab_quad$site)=="Kaguk Cove"]<-"Kaguk "
levels(hab_quad$site)[levels(hab_quad$site)=="Farallon Bay"]<-"Farallon"
levels(hab_quad$site)[levels(hab_quad$site)=="Chusini Cove"]<-"Chusini-Kladein Flat"
levels(hab_quad$site)[levels(hab_quad$site)=="South Fish Egg Island"]<-"South Fish Egg"
levels(hab_quad$site)[levels(hab_quad$site)=="Baker Island - kelp"]<-"Baker Island - kelp "
hab_quad$site <- as.character(hab_quad$site)

hab_quad2 <- left_join(hab_quad, loc, by = c("site" = "site_2019", "habitat" = "habitat")) %>%
  dplyr::select(-c(siteID_NOAA, site_2017, site_2018, place_name, study, latitude, longitude,
                   freshwater, sediment_description, general_description))
  
```


Write out files as csvs for upload to KNB (in their original format!)
```{r}
write.csv(hab_quad2, "Data/Habitat_data/hab_quadrat_2019.csv")
write.csv(hab_lng, "Data/Habitat_data/hab_length_2019.csv")
write.csv(hab_biom, "Data/Habitat_data/hab_species_biomass_2019.csv")
```

