---
title: "live_clams_and_assoc_eelgrass_cleaning_2022"
author: "Lia Domke"
date: "8/11/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In 2022 we were interested in understanding the competition between belowground biomass of eelgrass and live clams. To this end, we sampled clams at the 0 ft MMLW mark in appropriate clam habitat. We set a 100-m transect and used 0.25m^3 quadrats to sample 10 quadrats along the 100-m transect. This protocol was repeated from 2021 except 

At each of these quadrats we dug out the dirt, cleaned out the rocks/debris and non-clam items, and collected all live clams present, any algae, and any eelgrass (above and below ground shoots). We also recorded primary and secondary sediment type

This cleaning script will remove any misentries/spelling errors and combine the clam density with the eelgrass biomass. 

Read in libraries
```{r}
library(tidyverse)
```

read in data 
```{r}
live.clam <- read.csv("Data/clam_assoc_eelgrass_&density/clam_density_2022_RAW_8-12-22.csv", header = TRUE, stringsAsFactors = FALSE)
name <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A87744ced-530b-4bae-809e-ff1012c7ae72"),stringsAsFactors = FALSE, header = TRUE)
```


```{r}
unique(live.clam$IF.function) # okay so all the checking columns and entry columns match
glimpse(live.clam)
table(live.clam$size_mm == live.clam$size_mm_checked)

clam <- live.clam %>%
  dplyr::select(-c(IF.function, X, X.1, size_mm_checked)) %>%
  unite(bay_ID, bay_code:bay_sample) %>%
  mutate(common_name = tolower(common_name)) %>%
  mutate(species_code = ifelse(scientific_name == "Macoma inquinata", "CLMPOI", species_code)) %>%
  mutate(common_name = ifelse(scientific_name == "Macoma sp.", "macoma spp", common_name)) %>%
  mutate(species_code = ifelse(common_name == "macoma spp", "CLMMAC", species_code)) %>%
  mutate(scientific_name = ifelse(common_name == "macoma spp", "Macoma spp", scientific_name)) %>%
  add_column(year = 2022)
  
clam %>%
  dplyr::select(common_name:species_code) %>%
  distinct()

clam$size_mm <- as.numeric(clam$size_mm)
clam %>%
  group_by(bay_ID, quadrat_no, species_code, scientific_name, common_name) %>%
  dplyr::summarise(sum = n(), avg_mm = mean(size_mm))
  
```

```{r}
clam21 <- read.csv("Data/clam_assoc_eelgrass_&density/density_live_clams_from_live_clam_surveys_2021.csv")
names(clam21)
names(clam)

unique(clam21$site_name)
unique(clam$site_name)

head(name)
sites <- name %>%
  unite(bay_ID, bay_code:bay_sample) %>%
  dplyr::select(bay_ID, latitude, longitude)

live.clam21.22 <- clam21 %>%
  add_column(year = 2021) %>%
  dplyr::select(-X) %>%
  rbind(clam) %>%
  mutate(site_name = ifelse(bay_ID == "GUKT_A", "Guktu Bay", site_name)) %>%
  mutate(site_name = ifelse(bay_ID == "NATZ_A", "Natzuhini Bay", site_name)) %>%
  mutate(site_name = ifelse(bay_ID == "REFU_A", "Port Refugio", site_name)) %>%
  mutate(site_name = ifelse(bay_ID == "NOSK_A", "Nossuk Bay", site_name)) %>%
  left_join(sites, by = c("bay_ID")) %>%
  separate(bay_ID, into = c("bay_code", "bay_sample")) 

live.clam21.22 %>%
  dplyr::select(bay_code, bay_sample, site_name, latitude, longitude, year) %>%
  distinct()

# write it out
#write.csv(live.clam21.22, "Data/clam_assoc_eelgrass_&density/density_live_clams_from_live_clam_surveys_2021-2022.csv")
```
