---
title: "live_clams_and_assoc_eelgrass_cleaning_2021"
author: "Lia Domke"
date: "6/14/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---


In 2021 we were interested in understanding the competition between belowground biomass of eelgrass and live clams. To this end, we sampled clams and eelgrass at the 0 ft MMLW mark in appropriate clam habitat. We set a 100-m transect and used 0.25m^3 quadrats to sample 10 quadrats along the 100-m transect. 

At each of these quadrats we dug out the dirt, cleaned out the rocks/debris and non-clam items, and collected all live clams present, any algae, and any eelgrass (above and below ground shoots). We also recorded primary and secondary sediment type

This cleaning script will remove any misentries/spelling errors and combine the clam density with the eelgrass biomass. 

Read in libraries
```{r}
library(tidyverse)

```

Read in data
```{r}
assoc.eel <- read.csv("Data/clam_assoc_eelgrass_&density/clam_associated_eelgrass_algae_checked.csv", stringsAsFactors = FALSE, header = TRUE)
live.clam <- read.csv("Data/clam_assoc_eelgrass_&density/clam_density_2021_checked.csv", stringsAsFactors = FALSE, header = TRUE)

```


Data exploration
```{r}
glimpse(assoc.eel)

glimpse(live.clam) 


clam <- live.clam %>%
  dplyr::select(-c(X, X.1, X.2)) %>%
  unite(bay_ID, bay_code:bay_sample) %>%
  mutate(common_name = tolower(common_name)) %>%
  mutate(species_code = ifelse(common_name == "pointed macoma", "CLMPOI", species_code)) %>%
  mutate(common_name = ifelse(scientific_name == "Mya truncata", "truncate softshell", common_name)) %>%
  mutate(common_name = ifelse(species_code == "CLMMAC", "macoma spp", common_name)) %>%
  mutate(scientific_name = ifelse(species_code == "CLMMAC", "Macoma spp", scientific_name)) %>%
  mutate(bay_ID = ifelse(site_name == "North Fish Egg Island", "NFEI_A", bay_ID))


clam %>%
  dplyr::select(common_name:species_code) %>%
  distinct()

unique(clam$scientific_name)
unique(clam$species_code)

clam$size_mm <- as.numeric(clam$size_mm)
clam %>%
  group_by(bay_ID, quadrat_no, species_code, scientific_name, common_name) %>%
  dplyr::summarise(sum = n(), avg_mm = mean(size_mm))



# everything looks good for the live clams lets look at the eelgrass
```

looking at the associated eelgrass
```{r}
glimpse(assoc.eel)

eel <- assoc.eel %>%
  unite(bay_ID, bay_code:bay_sample)

unique(eel$bay_ID)
unique(eel$quadrat_no)
names(eel)
```
 
Things to do: 
- calculate dry mass (subtract the tin mass)
- calculate quadrat 2 that was processed in two batches
```{r}

cols <- names(dplyr::select(eel, belowground_empty_tin_mass_g:dry_algae_mass.tin_g))
eel[cols] <- lapply(eel[cols], as.numeric)

eel2 <-eel %>%
  mutate(quadrat_no = ifelse(quadrat_no == 2.1, 2.0, quadrat_no),
         quadrat_no = ifelse(quadrat_no == 2.2, 2.0, quadrat_no)) %>% # match up quadrat 2 in north fish egg 
  rowwise() %>%
  group_by(bay_ID, quadrat_no) %>% # sum up the two batches of processing for quadrat 2
  mutate(dry_belowground_mass.tin_g = sum(dry_belowground_mass.tin_g), 
         belowground_empty_tin_mass_g = sum(belowground_empty_tin_mass_g),
         dry_aboveground_mass.tin_g = sum(dry_belowground_mass.tin_g),
         aboveground_empty_tin_mass_g = sum(aboveground_empty_tin_mass_g),
         dry_algae_mass.tin_g = sum(dry_algae_mass.tin_g),
         algae_empty_tin_mass_g = sum(algae_empty_tin_mass_g),
         freshweight_aboveground_mass_g = sum(freshweight_aboveground_mass_g),
         freshweight_belowground_mass_g = sum(freshweight_belowground_mass_g),
         freshweight_algae_mass_g = sum(freshweight_algae_mass_g)) %>%
  distinct() %>%
  ungroup() %>% # calculate the actual mass (dry mass - tin mass)
  mutate(dry_belowground_eel_mass_g = dry_belowground_mass.tin_g - belowground_empty_tin_mass_g,
         dry_aboveground_eel_mass_g = dry_aboveground_mass.tin_g - aboveground_empty_tin_mass_g,
         dry_algae_mass_g = dry_algae_mass.tin_g - algae_empty_tin_mass_g) %>%
  mutate(notes = ifelse(notes == "quadrat 2 processed in two batches", "NA", notes)) # edit notes so it doesn't confuse people

# change na to 0 because the data was collected there just was nothing there
eel2[is.na(eel2)] <- 0
```

# combine the two data sets
```{r}
# we want to combine using the sites and quadrat_no

combined_clam_eel21 <- left_join(eel2, clam, by = c("bay_ID", "quadrat_no")) %>%
  dplyr::rename(notes = notes.y) %>%
  dplyr::select(bay_ID, site_name, sample_date, process_date, quadrat_no, quadrat_location_m, primary_sediment, secondary_sediment, belowground_empty_tin_mass_g:dry_algae_mass.tin_g, dry_belowground_eel_mass_g:samplers, notes)
  
```

The combined_clam_eel may or may not be what is wanted for final BCO-DMO submission, but its worth a shot. 

It could be best as separate csv - lets do that: 
```{r}
head(eel2) 
head(clam)
names(eel2)

# missing the place name nad long/lat for the eel2
sites <- read.csv("Data/Site_metadata_allyears_8-18-22.csv", stringsAsFactors = FALSE, header = TRUE) %>%
  unite(bay_id, bay_code:bay_sample) %>%
  dplyr::select(bay_id, place_name, latitude, longitude)

eel3 <- eel2 %>%
  left_join(sites, by = c("bay_ID" = "bay_id")) %>%
  separate(bay_ID, c("bay_code", "bay_sample"))
  
```

```{r}
write.csv(eel3, "Data/clam_assoc_eelgrass_&density/eelgrass&algae_associated_with_live_clam_surveys_2021.csv")
write.csv(clam, "Data/clam_assoc_eelgrass_&density/density_live_clams_from_live_clam_surveys_2021.csv")
```


```{r}
#write.csv(combined_clam_eel21, "Data/clam_assoc_eelgrass_&density/live_clam_associated_eel_combined_2021.csv")
```

