---
title: "seine_cleaning_script_2018"
author: "Lia Domke"
date: "1/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data cleaning for beach seine data

As apart of a project studying eelgrass ecosystems in southern Southeast Alaska, beach seines were done along the outer coast of Prince of Wales Island. A 37-m variable mesh net and a 16ft skiff was used to set a 'round haul' beach seine through eelgrass meadows. Fish were identified to the finest taxonomic level, enumerated, and measured either total length for most fish species or to fork length if the caudal fin was forked (e.g. salmon, shiner perch). After 30 measured fish, fish were no longer measured and lengths will be extrapolated from the 30 measured fish. 

This script is to clean the *dataset from 2018* of the eelgrass-associated fish community. The 2018 seines occured between June and July *only* and at a small subset of the previously sampled sites in 2017. 

This script will:

1. Classify invertebrates and vertebrates so that only fish can be filtered
2. Convert unmeasured fish to *measured* based on distributions of measured fish
3. Convert length data to *biomass* based on known relationships compiled from FishBase
4. Change site_code to new naming scheme that includes two column codes
5. Convert date to julian day and add year column

# Libraries

```{r libraries, include = FALSE}
library(dplyr)
library(tidyr)
library(vegan)
library(ggplot2)
library(lubridate)
```

# Data input

```{r data}
seine_all <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Aa595db78-ec1a-4783-b77e-43f90465bd83"), stringsAsFactors = FALSE, header = TRUE)

sp_names <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Abc823c8e-7be3-444b-a872-2e450ea3e85b"), stringsAsFactors = FALSE, header = TRUE)

site_names <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A87744ced-530b-4bae-809e-ff1012c7ae72"))

fishLW <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A42b3eee9-f6a3-4169-99ef-ed388f46d172"), stringsAsFactors = FALSE, header = TRUE) 
```


# Step 1. Filter by taxon
```{r}
unique(seine_all$species_common)
# 42 different species, need to remove non-fish species
# are all species present in the sp list?
anti_join(seine_all, sp_names, by = c("sp_code" = "SpCode"))

# remove non-fish species
fish <- seine_all %>% 
  filter(taxon == "Vertebrata")

unique(fish$taxon)

unique(fish$species_common) # only fish!
unique(fish$site) # still have all the sites! 
```
## Adjust sp_code values based on what came back from the 'anti_join' to match sp_names
```{r}
# Make columns factors in order to run code
seine_all$sp_code <- as.factor(seine_all$sp_code)
seine_all$species_common <- as.factor(seine_all$species_common)
seine_all$species_scientific <- as.factor(seine_all$species_scientific)

# Change tubenose poacher sp_code
levels(seine_all$sp_code)[levels(seine_all$sp_code)=="TUBEPOA"] <- "POATUBE"

# Change unknown snailfish sp_code, scientific name, and common name
levels(seine_all$species_common)[levels(seine_all$species_common)=="Snailfish"] <- "Juvenile snailfish"
levels(seine_all$species_scientific)[levels(seine_all$species_scientific)=="Liparis sp."] <- "Liparidae"
levels(seine_all$sp_code)[levels(seine_all$sp_code)=="SNAILFS"] <- "UNSNAILF"

# Adjust sp_code for masked greenling
levels(seine_all$sp_code)[levels(seine_all$sp_code)=="GREEMASK"] <- "GREENMAS"

anti_join(seine_all, sp_names, by = c("sp_code" = "SpCode"))
```

#Step 2. Unmeasured -> measured
This code comes from WR original script (Eelgrass_biometrics_data_analysis, L 325)

```{r}
#' From the raw data this script will export a file summarized by site ready for analysis. 
#' The goal will be to summaries this data at the site level. Summaries will occur in a few different
#' ways to prep the data for different types on analysis. There is 1 file generated here. 


#' All we need is the beach seine data and the fish length-mass conversions. 
#' make sure columns have appropriate character type
unique(fish$unmeasured_sm)

fish <- fish %>%
  dplyr::select(-c(unmeasured_sm)) # remove unmeasured_sm column because its empty


fish$unmeasured <- as.numeric(fish$unmeasured)
fish$fork_length <- as.numeric(fish$fork_length)

fish$abundance <- as.numeric(ifelse(is.na(fish$fork_length), paste(fish$unmeasured), 1))

df2 <- fish %>% # how many individual fish by site and species?
  group_by(site, sp_code) %>%
  dplyr::summarise(total = sum(abundance))

```

## Separate measured and unmeasured fish
```{r}
fish.m <- fish %>%
  # filter(taxon == "Vertebrata") %>%
  filter(fork_length != "NA")

fish.um <- fish %>%
  # filter(taxon == "Vertebrata") %>% 
  #filter(unmeasured != "estimate") %>% # assume that infield estimates are accurate
  filter(is.na(fork_length))

```

## Assign lengths to unmeasured fish

When beach seining we only measured the first 30 individuals of a species, and counted the rest. We can use the measured fishes to create a distribution from which we can assign 
lengths to the unmeasured fishes.

Assign lengths to unmeasured fish based on sampled distribution. This assignment should happen at the EventID level. i.e. use the distribution of fishes at a EventID to assign unmeasured fishes at that EventID. 

We will assume the least and just use the sampled proportions to assign lenghts to unmeasured fish. Exclued fishes that do not have a measured counterpart. These were insantces when we tossed a fish without knowing what it was other than it was a sculpin thing


Figure out which species at sites that there is not a measured conterpart 
```{r}
x <- fish.um %>%
  group_by(site, sp_code) %>%
  anti_join(fish.m, by = c("site", "sp_code")) # only 1 fish 

# want to remove these unmeasured fish from unmeasured df
fish.um.redu <- fish.um %>%
  group_by(site, sp_code) %>%
  anti_join(x, by = c("site", "sp_code"))

fish.um.redu$site <- as.factor(fish.um.redu$site)
# Now calculate a fork_length/total_length for each unmeasured fish based on the 
# distribution of the existing fish BY site (one site per seine).
```

## Run loop to fill unmeasured fish length to measured
```{r}
q <- data.frame() # empty dataframe to fill with for loop

for(s in unique(fish.um.redu$site)){ # cycle through unique sites (only one seine per site)
  m <- fish.m %>% # subset measured data by interation EventID
    filter(site == s)
  u <- fish.um.redu %>% # subset unmeasured data by iteration EventID
      filter(site == s)
  for(i in unique(u$sp_code)){ # cycle through species that are in UNMEASURED data
    samp <- m %>% # create sample from which to make distrubution
      filter(sp_code == i)
    unmeas <- u %>% # isolate unmeasured fish
      filter(sp_code == i)
    unmeas <- (as.numeric(sum(unmeas$unmeasured))) # save unmeasured value
    dat.temp1 <- data.frame(size = as.numeric(samp$fork_length))
    dat.temp2 <- dat.temp1 %>% 
      group_by(size) %>% 
      dplyr::summarize(count = n())
    dat.temp2$prob <- (dat.temp2$count/sum(dat.temp2$count))
    dat.temp2$x <- as.numeric(dat.temp2$size)
    fx <- function(n){ # function derived from limits and probabilities of above
    sample(x = (dat.temp2$x), size = n, replace = TRUE, prob = dat.temp2$prob)
    }
    dat.temp3 <- data.frame(site = s, sp_code = i, fork_length = fx(unmeas))
    q <- rbind(q, dat.temp3) # append iteration to full data
    q[] <- lapply(q, as.character)
  }
} # this is returning an error in regards to prob, but i am not sure what its talking about
```
## Loop check/add site info
Append assigned lengths to master data with all site level data

```{r}
# Extract site and sp data
fish.site <- unique(seine_all[,c("site", "date", "YYYYMMDD", "start_time", "end_time", "slope", "tide_height", "tide_time")])

fish.sp <- unique(fish[,c("species_common", "species_scientific", "sp_code", "taxon")])


## Merge with loop output, the lengths of unmeasured fishes ##
d.info <- left_join(q, fish.site, by = "site")

## Add species detail ##
d.info <- merge(d.info, fish.sp, by = "sp_code")

test <- anti_join(d.info, q, by = "sp_code")

## Merge with original measured fishes ##
d.info$fork_length <- as.numeric(d.info$fork_length)
fish.all <- fish.m %>%
  dplyr::select(site, date, YYYYMMDD, start_time, end_time, slope, tide_height,
                tide_time, species_common, species_scientific, sp_code, fork_length, 
                taxon, sex, notes) %>%
  bind_rows(d.info)

# Do some basic checks to make sure the the conversion from unmeasured to measured worked correctly
unique(fish.all$site) # 8 sites, correct
unique(fish.all$sp_code) # 33 levels of fish species
unique(fish$sp_code) # 33 levels of fish species in the original dataframe, correct! 7 unknown sp. 

# count total number of fish in original dataframe
fish$abundance <- as.numeric(ifelse(is.na(fish$fork_length), paste(fish$unmeasured), 1)) 

fish %>%
  dplyr::summarise(counts = sum(abundance)) # 11038

# does this match the fish.all dataframe?
fish.all$abundance <- as.numeric(ifelse(is.na(fish.all$fork_length), paste(fish.all$unmeasured), 1)) 

fish.all %>%
  dplyr::summarise(counts = sum(abundance)) # 11032
```

# Step 3. Convert length to biomass
## Calculate Biomass
Using the length-weight conversion values individual lengths will be converted to biomass. Coefficients are in cm*g so fork lengths will need to be converted to cm from mm.

First the L-W conversion data will need to be prepped for use. This will include defineing one a value (collapse the a TL column), define sp_code for new species, and filter out estimates that were made from SL (standard length) and fish calssified as UNFISH or unidentified fish, and then summarise by taking the average a and b value for each species.


## Put parameters into one column 
```{r}
fishLW$a_cm.g <- ifelse(is.na(fishLW$a_cm.g), fishLW$aTL_cm.g, fishLW$a_cm.g)

## define sp_code for new species ##
# Shorthorn sculpin #
fishLW$sp_code <- ifelse(fishLW$species_common == "Shorthorn sculpin", paste("SCULSHRN"), fishLW$sp_code)

# Longhorn sculpin #
fishLW$sp_code <- ifelse(fishLW$species_common == "Longhorn sculpin", paste("SCULLHRN"), fishLW$sp_code)

# Rockfish general #
fishLW$sp_code <- ifelse(fishLW$species_common == "Rockfish", paste("ROCKSEB"), fishLW$sp_code)

## Filter and summarise ##
LW <- fishLW %>%
  filter(Type != "SL") %>% 
  filter(sp_code != "UNFISH") %>% 
  group_by(sp_code) %>% 
  dplyr::summarise(a = mean(a_cm.g),
            b = mean(b_cm.g))
```

We know that we do not have estimates for a and b estiamtes for all the species so we need to apply other values to other species. First identify what species we need to have stand ins for:

1. "UNSCUL" - Unidentified sculpins will use the average a and b values for all sculpins species.
2. "UNFLAT" - Unidentified flatfish will use the average a and b values for all flatfish species.
3. "UNGREEN" - Unidentified greenling will use the average a and b values for all greenling speices.
4. "UNMYOXO" - Unidentified Myoxocephalus sp. will use the the average of a and b values of sampled members of the Myoxocephalus genus.
5. "UNGUNN" - Unidentified gunnel will use average a and b values from the sampled gunnel species
6. "UNARTE" - Unidentified Aretedius sp. will use the average a nad b values of sampled members of the Artedeius genus.


## Unique species in master data
```{r}
sp <- data.frame(sp_code = unique(fish.all$sp_code))

## Species that need stand ins ##
nomatch <- anti_join(sp, LW, by = "sp_code")


## Summarising stand ins ##
# Sculpins
scul <- data.frame(sp_code = "UNSCUL", fishLW %>% 
        filter(sp_code == "SCULBUF" | sp_code == "SCULGRT" | sp_code == "SCULGRU" |
                 sp_code == "SCULLHRN" | sp_code == "SCULMAN" |sp_code == "SCULPAD" | 
                 sp_code == "SCULPSTG" | sp_code == "SCULSAIL" | sp_code == "SCULSHRN" | 
                 sp_code =="SCULSILV" | sp_code == "SCULSMO" | sp_code == "SCULTIDE") %>% 
        summarise(a = mean(a_cm.g),
                  b = mean(b_cm.g)))

# Flats
flat <- data.frame(sp_code = "UNFLAT", fishLW %>% 
                     filter(sp_code == "SDABPAC" | sp_code == "SDABSPKL" | sp_code == "SOLEBUT" |
                              sp_code == "SOLECO" | sp_code == "SOLEENG" |sp_code == "SOLEROC" | 
                              sp_code == "FLOUNST") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

# Greenlings
grns <- data.frame(sp_code = "UNGREEN", fishLW %>% 
                     filter(sp_code == "GREENKEL" | sp_code == "GREENMAS" | sp_code == "GREENWHI") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

# Myoxocephalus
myoxo <- data.frame(sp_code = "UNMYOXO", fishLW %>% 
                      filter(sp_code == "SCULGRT" | sp_code == "SCULLHRN" | sp_code == "SCULSHRN") %>% 
                      summarise(a = mean(a_cm.g),
                                b = mean(b_cm.g)))

# Gunnels
gunn <- data.frame(sp_code = "UNGUNN", fishLW %>% 
                     filter(sp_code == "GUNNCRE" | sp_code == "GUNNPEN") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

arte <- data.frame(sp_code = "UNARTE", fishLW %>% 
                     filter(sp_code == "SCULPAD" | sp_code == "SCULSMO") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

## Master a b lsit ##
LW.master <- rbind.data.frame(LW, scul, flat, grns, myoxo, gunn, arte)

## Filter out UNFISH from fish.all ##
fish.all <- fish.all %>% 
  filter(sp_code != "UNFISH")

## Check that everything is cool ##
anti_join(fish.all, LW.master, by = "sp_code")

# Finally join the a and b values from above with the master data and calcualte biomass.
## Merge a and b values ##
fish.all.m <- merge(fish.all, LW.master, by = "sp_code")

## Convert mm to cm ##
fish.all.m$fork_length_cm <- fish.all.m$fork_length / 10

## Calculate mass in g ##
fish.all.m$mass_g <- (fish.all.m$a * fish.all.m$fork_length_cm^fish.all.m$b)

## some plots, cuz ya know ##
hist(fish.all.m$mass_g)
hist(log(fish.all.m$mass_g))
range(fish.all.m$mass_g)

## Clean up ##
fish.all.m$taxon[is.na(fish.all.m$taxon)] <- "Vertebrata"

# remove unused rows and readd in tide height
# extra tide height from original df
tide_h <- fish %>%
  group_by(site) %>%
  dplyr::select(site, tide_height) %>%
  distinct(tide_height)

fish.all.m <- fish.all.m %>%
  left_join(tide_h, by = "site") %>%
  mutate(tide_height = tide_height.x) %>%
  dplyr::select(-c(sex, abundance, tide_height.y, tide_height.x))
  

## Checking ##
fish.site.g <- fish.all.m %>% 
  group_by(site) %>% 
  dplyr::summarise(site_mass_g = sum(mass_g))

fish.site.g$mass_kg <- fish.site.g$site_mass_g / 1000

boxplot(fish.site.g$site_mass_g)

```

# Step 4. Convert date to JD and add year column
```{r}
fish.all.m <- fish.all.m %>% 
  mutate(date = ymd(YYYYMMDD)) %>% # convert to date format
  mutate(julian = yday(date)) %>% # make the date a julian day
  mutate(year = year(date)) # make a new column with just year
```

# Step 5. Convert site names to 2-column code system
```{r}
fish.all.m <- fish.all.m %>% 
  left_join(site_names, by = c("site" = "site_2018")) %>% # this needs to be changed for e/a YEAR. 
  select(-c(site_2017, freshwater, general_description, sediment_description, siteID_NOAA, study, 
            site_2019, a, b, fork_length_cm, site, YYYYMMDD, notes))

names(fish.all.m)
# place name gives the physical location, bay_code and bay_sample together give you each
# unique site that was sampled. 
```

# Export
```{r}
#write.csv(fish.all.m, "../Data/fish_mass_2018_derived.csv", row.names = FALSE)
```


