---
title: "seine_cleaning_2021"
author: "Lia Domke"
date: "9/13/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data cleaning for beach seine data

As apart of a project studying eelgrass ecosystems in southern Southeast Alaska, beach seines were done along the outer coast of Prince of Wales Island and in the vicinity of Ketchikan (Gravina Island and Annette Island). A 37-m variable mesh net and a 18ft skiff was used to set a 'round haul' beach seine through eelgrass meadows and understory kelp. Fish were identified to the finest taxonomic level, enumerated, and measured either total length for most fish species or to fork length if the caudal fin was forked (e.g. salmon, shiner perch). After 30 measured fish, fish were no longer measured and lengths will be extrapolated from the 30 measured fish. For some fish (like sticklebacks and shiner perch) were there were TWO DISTINCT size classes (juveniles and adults) 30 fish from each approximate age class were measured and fish counted but unmeasured were classified as *large* or *small* in the notes section of the csv files or the counts included in a seperate column (unmeasured_sm). 

This script is to clean the *dataset from 2021* of the eelgrass-associated and understory kelp associated fish community. The 2021 seines occured between June, July, and August and happened at sites that were previously seined in the 1990s by NMFS and 6 monitoring sites that have been seined yearly in July since 2020 (and seined 2017-2019 too just not in July). These sites were previously seined as part of the NOAA Nearshore Atlas of Fishes study that occured all throughout Alaska between 1998-2011. See  Johnson et al., 2012 or [here](https://www.fisheries.noaa.gov/alaska/habitat-conservation/nearshore-fish-atlas-alaska) for more information

This script will:

1. Data checking 2021 seines and adjusting dataframe appropriately
2. Classify invertebrates and vertebrates so that only fish can be filtered
3. Convert unmeasured fish to *length measurements*
4. Convert length data to *biomass*
5. Convert date to julian day and add year column
6. Change site_code to new naming scheme that includes two column codes

```{r libraries}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
```

Read in data
```{r data}
# 2021 seine data
# eventually replace with knb url when the raw data file is uploaded
seine21 <- read.csv("../2021_data/fish_beach_seine_2021_MASTER_9-3-21.csv", stringsAsFactors = FALSE, header = TRUE)

# species names list
sp_names <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Abc823c8e-7be3-444b-a872-2e450ea3e85b"), stringsAsFactors = FALSE, header = TRUE)

# site names (from knb 9/14/21 - doesn't have up to date sites)
site_names <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Ac9c99ce9-fbdd-4879-a2c9-c90448cdba7b"), stringsAsFactors = FALSE, header = TRUE)

# fish length to width (use this local csv because the updated LW w/ additional species isn't uploaded to knb yet)
fishLW <- read.csv("Data/fish_length_weight_conversions_cleaned_2021.csv", stringsAsFactors = FALSE, header = TRUE) 
```

# Step 1. Data checking

## Basic data checking
```{r}
glimpse(seine21)

# Look for data entry mistake where there are multiple dates associated with one site/sampling event
unique(seine21[,c("place_name", "habitat", "date", "YYYYMMDD", "start_time", "end_time", "slope", "tide_height", "tide_time")]) # all good

# Check for any mistakes with species code entry or site entry
unique(seine21$sp_code) # 64 unique species - including invertebrates
unique(seine21$bay_code) # north fish egg island bay code is wrong (should be NFEI)

# change bay code
seine21$bay_code <- as.factor(seine21$bay_code) 
levels(seine21$bay_code)[levels(seine21$bay_code)=="NEFI"] <- "NFEI"
seine21$bay_code <- as.character(seine21$bay_code)
unique(seine21$bay_code) # fixed

# Change several species code entries so that we can make length to width 
# Make species code a factor first
seine21$sp_code <- as.factor(seine21$sp_code)

# change SCULSCA scalyhead sculpin, should be classified with Artedius sp.
# This is because fish within the Artedius genus are hard to classify to species level. 
# Sometimes 
levels(seine21$sp_code)[levels(seine21$sp_code)=="SCULSCA"] <- "UNARTE"

# roughback got entered with two species codes SCULRBCK -> SCULRGH
levels(seine21$sp_code)[levels(seine21$sp_code)=="SCULRGH"] <- "SCULRBCK"

# Also have to repeat this for the fishLW dataframe
fishLW$sp_code <- as.factor(fishLW$sp_code)
levels(fishLW$sp_code)[levels(fishLW$sp_code)=="SCULRGH"] <- "SCULRBCK"


# change LORDBI -> UNLORD b/c difficult to impossible to differenciate from Red irish lord
#levels(seine$sp_code)[levels(seine$sp_code)=="LORDBI"] <- "UNLORD"

levels(seine21$sp_code)[levels(seine21$sp_code)=="UNSOLE"] <- "UNFLAT"
levels(seine21$species_common)[levels(seine21$species_common)=="Unidentified sole"] <- "Juvenile flatfish"

# change both sp_code columns back to characters
seine21$sp_code <- as.character(seine21$sp_code)
fishLW$sp_code <- as.character(fishLW$sp_code)

# check
unique(seine21$sp_code) #Now 63 instead of 64 species
anti_join(seine21, sp_names, by = c("sp_code" = "SpCode"))
# Note: 
seine <- seine21 %>%
  filter(taxon == "vertebrate") %>%
  unite(bay_ID, bay_code:bay_sample)
nonmatch1 <- anti_join(seine, fishLW, by = c("sp_code" = "sp_code"))
unique(nonmatch1$sp_code) # There a few more fish w/o length-weight relationship will be dealt with below.

# make sure there are no "Falses" in the data checking column (leftover from QAQC in entry)
unique(seine$same) # only true good
# remove QAQC columns
seine <- dplyr::select(seine, -c(FL_checking:same.))

# remove any trailing whitespaces
seine <- data.frame(lapply(seine, trimws), stringsAsFactors = FALSE)
```

## Seperate size classes
In 2021 there were fish--like shiner perch that included two different size classes (adults and juveniles). Therefore when there were more than 30 individuals of those species 30 of *each* size class were measured creating a bimodial distribution. When extrapolating lengths to the unmeasured fish it is important to use probabilities from the correct distribution. 

```{r}
# the unmeasured small were seperated from the large by seine location and species. So if a seine had both large and small individuals of a single species there is a number of unmeasured small in the 'unmeasured_sm' column. If not there wont be anything in that column just in 'unmeasured' (if more than 30 were caught)
unique(seine$notes)
unique(seine$unmeasured_sm)
```
# Size distributions
For the two species that were subset into large and small subset what are their size distributions/frequency?
Is there a break between the "juvenile" and "adult" size classes? 
Three-spine Stickleback and Shiner Perch
```{r}
seine$fork_length <- as.numeric(seine$fork_length)
# subset data by species
shiner <- seine %>%
  filter(sp_code == "PERCHSH")

stickle <- seine %>%
  filter(sp_code == "STICK3")

tubesnout <- seine %>%
  filter(sp_code == "TUBESNT")
```

Graph the size distributions of the measured fish
```{r}
# which sites had unmeasured small shiners 
shiner %>%
  filter(!is.na(unmeasured_sm)) # Bostwick and Nichols (BOST_A, NICH_A)

sh <- shiner %>%
  filter(bay_ID == "BOST_A" | bay_ID == "NICH_A") %>%
  ggplot() +
  geom_histogram(mapping = aes(x = fork_length), binwidth = 1)

sh + facet_wrap(~bay_ID)

#bimodal graph-- looks like the distributional cut off varies by site, so we'll have to take that into account when we run this... Bost-a cut off around 80 mm and for Nich-a ~ 100

stickle %>%
  filter(!is.na(unmeasured_sm)) # Refu-a, nefi-a, shin-a, natz-a

st <- stickle %>%
  filter(bay_ID == "REFU_A" | bay_ID == "NFEI_A" |bay_ID == "SHIN_A"| bay_ID == "NATZ_A") %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = fork_length)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))
st + facet_wrap(~bay_ID)
# bimodial again, looks like 45 mm and below is the small cut off for three spine stickleback  woo!

tubesnout %>%
  filter(!is.na(unmeasured_sm)) # there's only one site w/ small tubes

tubesnout %>%
  filter(bay_ID == "BOST_B") %>%
  ggplot() +
  geom_histogram(mapping = aes(x = fork_length)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))
# there were only a few large tubesnouts caught at this site and those were above ~ 100 mm 

```

### Fix taxon classification for trial seines
```{r}
# based on sp_code combine with seine data
#seine <- seine %>%
 # left_join(sp_names, by = c("sp_code" = "SpCode")) %>%
#  mutate(taxon = Taxa, sex = Sex) %>%
 # dplyr::select(bay_ID, place_name, date, YYYYMMDD, start_time, end_time, slope, tide_height, tide_time, 
 #               species_common, species_scientific, sp_code, fork_length, unmeasured, unmeasured_sm, 
  #              sex, notes, taxon)

seine$taxon <- as.factor(seine$taxon)
levels(seine$taxon)[levels(seine$taxon)=="fish"] <- "Vertebrata"
seine$taxon <- as.character(seine$taxon)
```

# Step 2. Classify taxons
```{r}
glimpse(seine)
unique(seine$species_common)
# 57 different species
fish <- seine

unique(fish$taxon)

unique(fish$species_common) # only fish!
unique(fish$bay_ID) # still have all the sites! 
```

#Step 3. Unmeasured -> measured
This code comes from WR original script (Eelgrass_biometrics_data_analysis, L 325)

```{r}
#' From the raw data this script will export a file summarized by site ready for analysis. 
#' The goal will be to summaries this data at the site level. Summaries will occur in a few different
#' ways to prep the data for different types on analysis. There is 1 file generated here. 


#' All we need is the beach seine data and the fish length-mass conversions. 
#' make sure columns have appropriate character type

fish$unmeasured <- as.numeric(fish$unmeasured)
fish$fork_length <- as.numeric(fish$fork_length)

fish$abundance <- as.numeric(ifelse(is.na(fish$fork_length), 
                                    ifelse(is.na(fish$unmeasured), paste(fish$unmeasured_sm), paste(fish$unmeasured)), 1))

df2 <- fish %>% # how many individual fish by site and species?
  dplyr::summarise(total = sum(abundance)) # 11987

df3 <- fish %>%
  group_by(bay_ID, sp_code) %>%
  dplyr::summarise(total = sum(abundance))
```

## Separate measured and unmeasured fish
### Loop for subset species
First run a similar loop as below to calculate estimated fork lengths for the small and large subset species. 
These include (for 2021) ONLY PERCHSH, STICK3, TUBESNT and PIPEFISH at a small subset of sites (vary by species)
### 1. List of subset sites
```{r}
site_shiner <- c("NICH_A", "BOST_A", "SHIN_A")
site_stick <- c("REFU_A", "NFEI_A", "SHIN_A", "NATZ_A")
site_tube <- c("BOST_B")
site_pipe <- c("BOST_B")
```

### 2. Need to subset data by species and sites were they are subsampled
```{r}
fish$unmeasured <- as.numeric(fish$unmeasured)
fish$unmeasured_sm <- as.numeric(fish$unmeasured_sm)

fish.stick <- fish %>%
  filter(sp_code == "STICK3")

fish.shiner <- fish %>%
  filter(sp_code == "PERCHSH")

fish.tube <- fish %>%
  filter(sp_code == "TUBESNT")

fish.pipe <- fish %>%
  filter(sp_code == "PIPEFISH")

```

### 3. subset by the sites where those species were subsampled
```{r}
fish.stick.sub <- subset(fish.stick, bay_ID %in% site_stick)
fish.stick.sub %>%
    filter(is.na(fork_length)) %>%
    summarise(sum(abundance)) # 1251 unmeasured sticks
fish.shiner.sub <- subset(fish.shiner, bay_ID %in% site_shiner)
fish.shiner.sub %>%
    filter(is.na(fork_length)) %>%
    summarise(sum(abundance)) # 1341 unmeasured shiners
fish.tube.sub <- subset(fish.tube, bay_ID %in% site_tube)
fish.tube.sub %>%
    filter(is.na(fork_length)) %>%
    summarise(sum(abundance)) # 32 unmeasured tubes
fish.pipe.sub <- subset(fish.pipe, bay_ID %in% site_pipe)
fish.pipe.sub %>%
    filter(is.na(fork_length)) %>%
    summarise(sum(abundance)) # 42 unmeasured sticks
# total fish that need to be measured 2666
```

### 4. cycle through loops (1 for different species)
Sticklebacks first: based on distributions of known fork length we set a cut off at 45 mm. Both large and small subset
```{r}
# Need dataframes by species, size, and measured v. unmeasured
sm.m.st <- fish.stick.sub %>%
  filter(fork_length != "NA") %>%
  filter(fork_length <= 45)
lg.m.st <- fish.stick.sub %>%
  filter(fork_length != "NA") %>%
  filter(fork_length > 45)
sm.um.st <- fish.stick.sub %>%
  filter(is.na(fork_length)) %>%
  filter(unmeasured_sm != "NA")
lg.um.st <- fish.stick.sub %>%
  filter(is.na(fork_length)) %>%
  filter(notes == "large stickles" | notes == "large stickle")
```

####Loop for stickleback
Loop for each size class to approximate measurements for unmeasured fish
```{r}
lg_stick <- data.frame() # empty dataframe to fill with for loop

for(s in unique(fish.stick.sub$bay_ID)){ # cycle through unique sites (only one seine per site)
  m <- lg.m.st %>% # subset measured data by interation EventID
    filter(bay_ID == s)
  u <- lg.um.st %>% # subset unmeasured data by iteration EventID
      filter(bay_ID == s)
  for(i in unique(fish.stick.sub$sp_code)){ # cycle through species that are in UNMEASURED data
    samp <- m %>% # create sample from which to make distrubution
      filter(sp_code == i)
    unmeas <- u %>% # isolate unmeasured fish
      filter(sp_code == i)
    unmeas <- (as.numeric(sum(unmeas$unmeasured))) # save unmeasured value
    dat.temp1 <- data.frame(size = as.numeric(samp$fork_length))
    dat.temp2 <- dat.temp1 %>% 
      group_by(size) %>% 
      dplyr::summarize(count = n())
    dat.temp2$prob <- (dat.temp2$count/sum(dat.temp2$count))
    dat.temp2$x <- as.numeric(dat.temp2$size)
    fx <- function(n){ # function derived from limits and probabilities of above
    sample(x = (dat.temp2$x), size = n, replace = TRUE, prob = dat.temp2$prob)
    }
    dat.temp3 <- data.frame(site = s, sp_code = i, fork_length = fx(unmeas))
    lg_stick <- rbind(lg_stick, dat.temp3) # append iteration to full data
    lg_stick[] <- lapply(lg_stick, as.character)
  }
} 

lg_stick %>%
  group_by(site) %>%
  dplyr::summarise(n())
```

RE DO FOR THE SMALL SUBSET - sticklebacks
```{r}
sm_stick <- data.frame() # empty dataframe to fill with for loop

for(s in unique(fish.stick.sub$bay_ID)){ # cycle through unique sites (only one seine per site)
  m <- sm.m.st %>% # subset measured data by interation EventID
    filter(bay_ID == s)
  u <- sm.um.st %>% # subset unmeasured data by iteration EventID
      filter(bay_ID == s)
  for(i in unique(fish.stick.sub$sp_code)){ # cycle through species that are in UNMEASURED data
    samp <- m %>% # create sample from which to make distrubution
      filter(sp_code == i)
    unmeas <- u %>% # isolate unmeasured fish
      filter(sp_code == i)
    unmeas <- (as.numeric(sum(unmeas$unmeasured_sm))) # save unmeasured value
    dat.temp1 <- data.frame(size = as.numeric(samp$fork_length))
    dat.temp2 <- dat.temp1 %>% 
      group_by(size) %>% 
      dplyr::summarize(count = n())
    dat.temp2$prob <- (dat.temp2$count/sum(dat.temp2$count))
    dat.temp2$x <- as.numeric(dat.temp2$size)
    fx <- function(n){ # function derived from limits and probabilities of above
    sample(x = (dat.temp2$x), size = n, replace = TRUE, prob = dat.temp2$prob)
    }
    dat.temp3 <- data.frame(site = s, sp_code = i, fork_length = fx(unmeas))
    sm_stick <- rbind(sm_stick, dat.temp3) # append iteration to full data
    sm_stick[] <- lapply(sm_stick, as.character)
  }
} 

sm_stick %>%
  group_by(site) %>%
  dplyr::summarise(n())
  

m_stick <- bind_rows(sm_stick, lg_stick) # 1251
m_stick %>%
  group_by(site) %>%
  dplyr::summarise(n())

m_stick$fork_length <- as.numeric(m_stick$fork_length)
```

Shiner perch: Bost-a cut off around 80 mm and for Nich-a ~ 100

####Loop for shiner perch 
(have to add in site ifelse condition because of different size limits at diff sites, fish grow..)
```{r}
# Do for small subset (there was no large subset that need to be estimated)
sm_shiner <- data.frame() # empty dataframe to fill with for loop

# There are no small shiners in SHIN_A (only large) - so remove that bayID
fish.sm.shiner <- fish.shiner.sub %>%
  filter(bay_ID != "SHIN_A")

for(s in unique(fish.sm.shiner$bay_ID)){ # cycle through unique sites (only one seine per site)
  m <- fish.sm.shiner %>% # subset measured data by interation EventID
    filter(bay_ID == s) %>%
    filter(!is.na(fork_length)) %>%
    filter(ifelse(bay_ID == "BOST_A", fork_length < 80, fork_length < 100))
  u <- fish.sm.shiner %>% # subset unmeasured data by iteration EventID
    filter(bay_ID == s) %>%
    filter(is.na(fork_length)) %>%
    filter(unmeasured_sm != "NA")
  for(i in unique(fish.sm.shiner$sp_code)){ # cycle through species that are in UNMEASURED data
    samp <- m %>% # create sample from which to make distrubution
      filter(sp_code == i)
    unmeas <- u %>% # isolate unmeasured fish
      filter(sp_code == i)
    unmeas <- (as.numeric(sum(unmeas$unmeasured_sm))) # save unmeasured value
    dat.temp1 <- data.frame(size = as.numeric(samp$fork_length))
    dat.temp2 <- dat.temp1 %>% 
      group_by(size) %>% 
      dplyr::summarize(count = n())
    dat.temp2$prob <- (dat.temp2$count/sum(dat.temp2$count))
    dat.temp2$x <- as.numeric(dat.temp2$size)
    fx <- function(n){ # function derived from limits and probabilities of above
    sample(x = (dat.temp2$x), size = n, replace = TRUE, prob = dat.temp2$prob)
    }
    dat.temp3 <- data.frame(site = s, sp_code = i, fork_length = fx(unmeas))
    dat.temp3$fork_length <- as.numeric(dat.temp3$fork_length)
    sm_shiner <- bind_rows(sm_shiner, dat.temp3) # append iteration to full data
    #sm_shiner[] <- lapply(sm_shiner, as.character)
  }
} 

sm_shiner %>%
  group_by(site) %>%
  dplyr::summarise(n())

```

Loop for large shiner
(have to add in site ifelse condition because of different size limits at diff sites, fish grow..)
```{r}
# Do for small subset (there was no large subset that need to be estimated)
lg_shiner <- data.frame() # empty dataframe to fill with for loop

fish.lg.shiner <- fish.shiner.sub %>%
  filter(bay_ID != "BOST_A") # only small shiners in this site

for(s in unique(fish.lg.shiner$bay_ID)){ # cycle through unique sites (only one seine per site)
  m <- fish.lg.shiner %>% # subset measured data by interation EventID
    filter(bay_ID == s) %>%
    filter(!is.na(fork_length)) %>%
    filter(ifelse(bay_ID == "NICH_A", fork_length > 100, fork_length > 80))
  u <- fish.lg.shiner %>% # subset unmeasured data by iteration EventID
    filter(bay_ID == s) %>%
    filter(is.na(fork_length)) %>%
    filter(unmeasured != "NA")
  for(i in unique(fish.lg.shiner$sp_code)){ # cycle through species that are in UNMEASURED data
    samp <- m %>% # create sample from which to make distrubution
      filter(sp_code == i)
    unmeas <- u %>% # isolate unmeasured fish
      filter(sp_code == i)
    unmeas <- (as.numeric(sum(unmeas$unmeasured))) # save unmeasured value
    dat.temp1 <- data.frame(size = as.numeric(samp$fork_length))
    dat.temp2 <- dat.temp1 %>% 
      group_by(size) %>% 
      dplyr::summarize(count = n())
    dat.temp2$prob <- (dat.temp2$count/sum(dat.temp2$count))
    dat.temp2$x <- as.numeric(dat.temp2$size)
    fx <- function(n){ # function derived from limits and probabilities of above
    sample(x = (dat.temp2$x), size = n, replace = TRUE, prob = dat.temp2$prob)
    }
    dat.temp3 <- data.frame(site = s, sp_code = i, fork_length = fx(unmeas))
    dat.temp3$fork_length <- as.numeric(dat.temp3$fork_length)
    lg_shiner <- bind_rows(lg_shiner, dat.temp3) # append iteration to full data
    #lg_shiner[] <- lapply(lg_shiner, as.character)
  }
} 

lg_shiner %>%
  group_by(site) %>%
  dplyr::summarise(n())

m_shiner <- bind_rows(sm_shiner, lg_shiner)

m_shiner %>%
  group_by(site) %>%
  dplyr::summarise(n())
```

####Loop for tubesnout 
(only one site and only small tubes below ~100m)
```{r}
# Do for small subset (there was no large subset that need to be estimated)
sm_tube <- data.frame() # empty dataframe to fill with for loop

for(s in unique(fish.tube.sub$bay_ID)){ # cycle through unique sites (only one seine per site)
  m <- fish.tube.sub %>% # subset measured data by interation EventID
    filter(bay_ID == s) %>%
    filter(!is.na(fork_length)) %>%
    filter(fork_length < 100)
  u <- fish.tube.sub %>% # subset unmeasured data by iteration EventID
    filter(bay_ID == s) %>%
    filter(is.na(fork_length)) %>%
    filter(unmeasured_sm != "NA")
  for(i in "TUBESNT"){ # cycle through species that are in UNMEASURED data
    samp <- m %>% # create sample from which to make distrubution
      filter(sp_code == i)
    unmeas <- u %>% # isolate unmeasured fish
      filter(sp_code == i)
    unmeas <- (as.numeric(sum(unmeas$unmeasured_sm))) # save unmeasured value
    dat.temp1 <- data.frame(size = as.numeric(samp$fork_length))
    dat.temp2 <- dat.temp1 %>% 
      group_by(size) %>% 
      dplyr::summarize(count = n())
    dat.temp2$prob <- (dat.temp2$count/sum(dat.temp2$count))
    dat.temp2$x <- as.numeric(dat.temp2$size)
    fx <- function(n){ # function derived from limits and probabilities of above
    sample(x = (dat.temp2$x), size = n, replace = TRUE, prob = dat.temp2$prob)
    }
    dat.temp3 <- data.frame(site = s, sp_code = i, fork_length = fx(unmeas))
    sm_tube <- rbind(sm_tube, dat.temp3) # append iteration to full data
    sm_tube[] <- lapply(sm_tube, as.character)
  }
} 
sm_tube %>%
  group_by(site) %>%
  dplyr::summarise(n())

m_tube <- sm_tube
m_tube$fork_length <- as.numeric(m_tube$fork_length)
```
####Loop for pipes
 Do for small subset (there was no large subset that need to be estimated)
```{r}
sm_pipe <- data.frame() # empty dataframe to fill with for loop

for(s in unique(fish.pipe.sub$bay_ID)){ # cycle through unique sites (only one seine per site)
  m <- fish.pipe.sub %>% # subset measured data by interation EventID
    filter(bay_ID == s) %>%
    filter(!is.na(fork_length)) %>%
    filter(fork_length < 100)
  u <- fish.pipe.sub %>% # subset unmeasured data by iteration EventID
    filter(bay_ID == s) %>%
    filter(is.na(fork_length))
  for(i in "PIPEFISH"){ # cycle through species that are in UNMEASURED data
    samp <- m %>% # create sample from which to make distrubution
      filter(sp_code == i)
    unmeas <- u %>% # isolate unmeasured fish
      filter(sp_code == i)
    unmeas <- (as.numeric(sum(unmeas$unmeasured_sm))) # save unmeasured value
    dat.temp1 <- data.frame(size = as.numeric(samp$fork_length))
    dat.temp2 <- dat.temp1 %>% 
      group_by(size) %>% 
      dplyr::summarize(count = n())
    dat.temp2$prob <- (dat.temp2$count/sum(dat.temp2$count))
    dat.temp2$x <- as.numeric(dat.temp2$size)
    fx <- function(n){ # function derived from limits and probabilities of above
    sample(x = (dat.temp2$x), size = n, replace = TRUE, prob = dat.temp2$prob)
    }
    dat.temp3 <- data.frame(site = s, sp_code = i, fork_length = fx(unmeas))
    sm_pipe <- rbind(sm_pipe, dat.temp3) # append iteration to full data
    sm_pipe[] <- lapply(sm_pipe, as.character)
  }
} 
sm_pipe %>%
  group_by(site) %>%
  dplyr::summarise(n())

m_pipe <- sm_pipe
m_pipe$fork_length <- as.numeric(m_pipe$fork_length)
```

###Loop for rest
Subset the rest of the data into m and um 
```{r}
# need to filter out the small and large subsets for each species
# how many fish are there total
fish %>%
  summarise(sum(abundance)) # 11987

# remove the already extrapolated "unmeasured" species (done above)
remove_fish <- bind_rows(fish.shiner.sub, fish.stick.sub, fish.tube.sub, fish.pipe.sub) %>%
  filter(is.na(fork_length)) 

remove_fish %>%
  dplyr::summarise(sum(abundance)) # 2666 fish extrapolated above

fish.sub <- fish %>%
  dplyr::select(-notes) %>% # remove notes otherwise the antijoin won't work correctly
  anti_join(remove_fish) # joins by whatever ISN'T in remove_fish

# are there the right number of fish? (11987-2666 = 9321)
fish.sub %>%
  summarise(sum(abundance)) # yes! 

# seperate by umeasured and measured
fish.m <- fish.sub %>%
  filter(fork_length != "NA")

fish.um <- fish.sub %>%
  filter(is.na(fork_length))
```


## Assign lengths to unmeasured fish for other species

When beach seining we only measured the first 30 individuals of a species, and counted the rest. We can use the measured fishes to create a distribution from which we
can assign lengths to the unmeasured fishes.

Assign lengths to unmeasured fish based on sampled distribution. This assignment should happen at the EventID level. i.e. use the distribution of fishes at a 
EventID to assign unmeasured fishes at that EventID. 

We will assume the least and just use the sampled proportions to assign lenghts to unmeasured fish. Exclued fishes that do not have a measured counterpart. 
These were insantces when we tossed a fish without knowing what it was other than it was a sculpin thing


Figure out which species at sites that there is not a measured conterpart 
```{r}
x <- fish.um %>%
  group_by(bay_ID, sp_code) %>%
  anti_join(fish.m, by = c("bay_ID", "sp_code")) # none!

fish.um %>%
  dplyr::summarise(sum(unmeasured))

# Now calculate a fork_length/total_length for each unmeasured fish based on the 
# distribution of the existing fish BY site (one site per seine).
```

## Run loop to fill unmeasured fish length to measured
```{r}
w <- data.frame() # empty dataframe to fill with for loop

for(s in unique(fish.um$bay_ID)){ # cycle through unique sites (only one seine per site)
  m <- fish.m %>% # subset measured data by interation EventID
    filter(bay_ID == s)
  u <- fish.um %>% # subset unmeasured data by iteration EventID
      filter(bay_ID == s)
  for(i in unique(u$sp_code)){ # cycle through species that are in UNMEASURED data
    samp <- m %>% # create sample from which to make distrubution
      filter(sp_code == i)
    unmeas <- u %>% # isolate unmeasured fish
      filter(sp_code == i)
    unmeas <- (as.numeric(sum(unmeas$unmeasured))) # save unmeasured value
    dat.temp1 <- data.frame(size = as.numeric(samp$fork_length))
    dat.temp2 <- dat.temp1 %>% 
      group_by(size) %>% 
      dplyr::summarize(count = n())
    dat.temp2$prob <- (dat.temp2$count/sum(dat.temp2$count))
    dat.temp2$x <- as.numeric(dat.temp2$size)
    fx <- function(n){ # function derived from limits and probabilities of above
    sample(x = (dat.temp2$x), size = n, replace = TRUE, prob = dat.temp2$prob)
    }
    dat.temp3 <- data.frame(site = s, sp_code = i, fork_length = fx(unmeas))
    w <- rbind(w, dat.temp3) # append iteration to full data
    w[] <- lapply(w, as.character)
  }
}

w %>%
  dplyr::summarise(n()) # thats the same as above, loop worked

# Bind the now measured fish with fish from above (stick and shiner and tube and pipe)
w$fork_length <- as.numeric(w$fork_length)
all.unmeasured <- bind_rows(w, m_shiner, m_stick, m_tube, m_pipe)

all.unmeasured %>%
  dplyr::summarise(n())
```

## Loop check/add site info
Append assigned lengths to master data with all site level data

```{r}
# Extract site and sp data
fish.site <- unique(seine[,c("bay_ID", "date", "YYYYMMDD", "start_time", "end_time", "slope", "tide_height", "tide_time")])

seine$species_common <- trimws(seine$species_common, "r") # remove hanging white space on right side
fish.sp <- unique(seine[,c("species_common", "species_scientific", "sp_code", "taxon")])
# need to remove duplicate case of UNARTE - both scalyhead and juvenile artedius get classified as UNARTE for species code, so when you join by sp_code (below) it artificially duplicates the number of UNARTE
fish.sp <- fish.sp %>%
  filter(species_common != "Scalyhead sculpin")

## Merge with loop output, the lengths of unmeasured fishes ##
d.info <- left_join(all.unmeasured, fish.site, by = c("site"= "bay_ID")) # but this dropped 2 sites because there were two sites w/o any unmeasured fishes

## Add species detail ##
d.info <- d.info %>%
  left_join(fish.sp, by = "sp_code") %>%
  dplyr::rename(bay_ID = site)

## were any species left out?
anti_join(d.info, test, by = "sp_code") # none

## Merge with original *measured* fishes ##
d.info$fork_length <- as.numeric(d.info$fork_length)
fish.all <- seine %>%
  filter(fork_length != "NA") %>%
  dplyr::select(bay_ID, date, YYYYMMDD, start_time, end_time, slope, tide_height,
                tide_time, species_common, species_scientific, sp_code, fork_length, 
                taxon) %>%
  bind_rows(d.info)

# Do some basic checks to make sure the the conversion from unmeasured to measured worked correctly
unique(fish.all$bay_ID) # 22 sites
unique(fish.all$sp_code) # 56 levels of fish species
unique(fish$sp_code) # 56 levels of fish species in the original dataframe, correct! 

# count total number of fish in original dataframe
seine$abundance <- as.numeric(ifelse(is.na(seine$fork_length), 
                                    ifelse(is.na(seine$unmeasured), paste(seine$unmeasured_sm), paste(seine$unmeasured)), 1))
seine %>%
  dplyr::summarise(counts = sum(abundance)) # 11987

# count by site and species
df3 <- seine %>%
  group_by(bay_ID, sp_code) %>%
  dplyr::summarise(total = sum(abundance))

# does this match the new fish.all dataframe?
fish.all$abundance <- as.numeric(ifelse(is.na(fish.all$fork_length), paste(fish.all$unmeasured), 1)) 

fish.all %>%
  dplyr::summarise(counts = sum(abundance)) #11987, yes!

df4 <- fish.all %>%
  group_by(bay_ID, sp_code) %>%
  dplyr::summarise(counts = sum(abundance))

# check w/ previous numbers (df3) If there are issues with more individuals of e/a species after the forloop to 
# convert unmeasured fish to measured fish make sure to check the extracted site/date info. When you merge the loop
# output with the site info if there are duplicated sites, it will duplicate the number of fish--artificially 
# inflating the number of fish at ea site. 
check_fish_no <- merge(df4, df3)
check_fish_no$same <- ifelse(df4$counts == df3$total, TRUE, FALSE)
unique(check_fish_no$same) # no false!

```

# Step 4. Convert length to biomass
## Calculate Biomass
Using the length-weight conversion values individual lengths will be converted to biomass. Coefficients are in cm*g so fork lengths will need to be converted to cm from mm.

First the L-W conversion data will need to be prepped for use. This will include defineing one a value (collapse the a TL column), define sp_code for new species, and filter out estimates that were made from SL (standard length) and fish calssified as UNFISH or unidentified fish, and then summarise by taking the average a and b value for each species.


## Put parameters into one column 
```{r}
fishLW$a_cm.g <- ifelse(is.na(fishLW$a_cm.g), fishLW$aTL_cm.g, fishLW$a_cm.g)

## define sp_code for new species ##
# Shorthorn sculpin #
fishLW$sp_code <- ifelse(fishLW$species_common == "Shorthorn sculpin", paste("SCULSHRN"), fishLW$sp_code)

# Longhorn sculpin #
fishLW$sp_code <- ifelse(fishLW$species_common == "Longhorn sculpin", paste("SCULLHRN"), fishLW$sp_code)

# Rockfish general #
fishLW$sp_code <- ifelse(fishLW$species_common == "Rockfish", paste("ROCKSEB"), fishLW$sp_code)

## Filter and summarise ##
LW <- fishLW %>%
  filter(Type != "SL") %>% 
  filter(sp_code != "UNFISH") %>% 
  group_by(sp_code) %>% 
  dplyr::summarise(a = mean(a_cm.g),
            b = mean(b_cm.g))
```

We know that we do not have estimates for a and b estiamtes for all the species so we need to apply other values to other species. First identify what species we need to have stand ins for:

1. "UNSCUL" - Unidentified sculpins will use the average a and b values for all sculpins species.
2. "UNFLAT" - Unidentified flatfish will use the average a and b values for all flatfish species.
3. "UNGREEN" - Unidentified greenling will use the average a and b values for all greenling speices.
4. "UNMYOXO" - Unidentified Myoxocephalus sp. will use the the average of a and b values of sampled members of the Myoxocephalus genus.
5. "UNGUNN" - Unidentified gunnel will use average a and b values from the sampled gunnel species
6. "UNARTE" - Unidentified Aretedius sp. will use the average a nad b values of sampled members of the Artedeius genus.
7. "UNSNAIL" - Unidentified Liparis sp. will use the average a nad b values of sampled members of the Liparidae genus.

## Unique species in master data
If there isn't already alpha and beta values for each species required can use *fish_lng_bms_cleaning.R* script to add in info. Most of the values are from fishbase
```{r}
sp <- data.frame(sp_code = unique(fish.all$sp_code))

## Species that need stand ins ##
nomatch <- anti_join(sp, LW, by = "sp_code") 


## Summarising stand ins ##
# Sculpins
scul <- data.frame(sp_code = "UNSCUL", fishLW %>% 
        filter(sp_code == "SCULBUF" | sp_code == "SCULGRT" | sp_code == "SCULGRU" |
                 sp_code == "SCULLHRN" | sp_code == "SCULMAN" |sp_code == "SCULPAD" | 
                 sp_code == "SCULPSTG" | sp_code == "SCULSAIL" | sp_code == "SCULSHRN" | 
                 sp_code =="SCULSILV" | sp_code == "SCULSMO" | sp_code == "SCULTIDE") %>% 
        summarise(a = mean(a_cm.g),
                  b = mean(b_cm.g)))

# Flats
flat <- data.frame(sp_code = "UNFLAT", fishLW %>% 
                     filter(sp_code == "SDABPAC" | sp_code == "SDABSPKL" | sp_code == "SOLEBUT" |
                              sp_code == "SOLECO" | sp_code == "SOLEENG" |sp_code == "SOLEROC" | 
                              sp_code == "FLOUNST" | sp_code == "SOLEYEL") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

# Greenlings
grns <- data.frame(sp_code = "UNGREEN", fishLW %>% 
                     filter(sp_code == "GREENKEL" | sp_code == "GREENMAS" | sp_code == "GREENWHI" |
                              sp_code == "GREENPAI") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

# Myoxocephalus
myoxo <- data.frame(sp_code = "UNMYOXO", fishLW %>% 
                      filter(sp_code == "SCULGRT" | sp_code == "SCULLHRN" | sp_code == "SCULSHRN") %>% 
                      summarise(a = mean(a_cm.g),
                                b = mean(b_cm.g)))

# Gunnels
gunn <- data.frame(sp_code = "UNGUNN", fishLW %>% 
                     filter(sp_code == "GUNNCRE" | sp_code == "GUNNPEN") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

# Artedius
arte <- data.frame(sp_code = "UNARTE", fishLW %>% 
                     filter(sp_code == "SCULPAD" | sp_code == "SCULSMO") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))
# Rockfish
rock <- data.frame(sp_code = "UNROCK", fishLW %>% 
                     filter(sp_code == "ROCKSEB" | sp_code == "ROCKSLG"
                            | sp_code == "ROCKCOP" | sp_code == "ROCKQUI"
                            | sp_code == "ROCKBLA" | sp_code == "ROCKBRO") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

rock2 <- data.frame(sp_code = "ROCKSP", fishLW %>% 
                     filter(sp_code == "ROCKSEB" | sp_code == "ROCKSLG"
                            | sp_code == "ROCKCOP" | sp_code == "ROCKQUI"
                            | sp_code == "ROCKBLA" | sp_code == "ROCKBRO") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

# Snailfish
snail <- data.frame(sp_code = "UNSNAILF", fishLW %>% 
                     filter(sp_code == "SNAILFR" | sp_code == "SNAILFTI"
                            | sp_code == "SNAILFSH" | sp_code == "SNAILFRT") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

# Irish Lord
lord <- data.frame(sp_code = "UNLORD", fishLW %>% 
                     filter(sp_code == "LORDBI" | sp_code == "LORDRI") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

## Master a b lsit ##
LW.master <- rbind.data.frame(LW, scul, flat, grns, myoxo, gunn, arte, rock, snail, lord, rock2)

## Filter out UNFISH from fish.all ##
fish.all <- fish.all %>% 
  filter(sp_code != "UNFISH") %>%
  filter(sp_code != "UNLARV")

## Check that everything is cool ##
unique(anti_join(fish.all, LW.master, by = "sp_code")$sp_code) # no unmatched sp_codes! awesome
```

# Step 5. Convert date to JD and add year column
```{r}
fish.all.m <- fish.all.m %>% 
  mutate(date = ymd(YYYYMMDD)) %>% # convert to date format
  mutate(julian = yday(date)) %>% # make the date a julian day
  mutate(year = year(date)) # make a new column with just year
```

# Step 6. Convert site names to 2-column code system
```{r}
fish.all.m <- fish.all.m %>% 
  left_join(site_names, by = c("site" = "site_2019")) %>% # this needs to be changed for e/a YEAR. 
  select(-c(site_2017, site_2018, freshwater, general_description, sediment_description, siteID_NOAA, study, notes, fork_length_cm, site, YYYYMMDD, a, b))

names(fish.all.m)
# place name gives the physical location, bay_code and bay_sample together give you each
# unique site that was sampled. 

```
```{r}
# Finally join the a and b values from above with the master data and calcualte biomass.
## Merge a and b values ##
fish.all.m <- merge(fish.all, LW.master, by = "sp_code")

## Convert mm to cm ##
fish.all.m$fork_length_cm <- fish.all.m$fork_length / 10

## Calculate mass in g, a*FL^b ##
fish.all.m$mass_g <- (fish.all.m$a * fish.all.m$fork_length_cm^fish.all.m$b)

## some plots, cuz ya know ##
hist(fish.all.m$mass_g)
hist(log(fish.all.m$mass_g))
range(fish.all.m$mass_g)

## Clean up ##
fish.all.m$taxon[is.na(fish.all.m$taxon)] <- "Vertebrata"

# remove unused rows and readd in tide height
# extra tide height from original df
tide_h <- fish %>%
  group_by(site) %>%
  dplyr::select(site, tide_height) %>%
  distinct(tide_height)

fish.all.m <- fish.all.m %>%
  left_join(tide_h, by = "site") %>%
  mutate(tide_height = tide_height.x) %>%
  dplyr::select(-c(sex, abundance, tide_height.y, tide_height.x))
  

## Checking ##
fish.site.g <- fish.all.m %>% 
  group_by(site) %>% 
  dplyr::summarise(site_mass_g = sum(mass_g))

fish.site.g$mass_kg <- fish.site.g$site_mass_g / 1000

boxplot(fish.site.g$site_mass_g)

```