---
title: "seine_cleaning_2020"
author: "Lia Domke"
date: "1/24/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data cleaning for beach seine data

As apart of a project studying eelgrass ecosystems in southern Southeast Alaska, beach seines were done along the outer coast of Prince of Wales Island. A 37-m variable mesh net and a 16ft skiff was used to set a 'round haul' beach seine through eelgrass meadows. Fish were identified to the finest taxonomic level, enumerated, and measured either total length for most fish species or to fork length if the caudal fin was forked (e.g. salmon, shiner perch). After 30 measured fish, fish were no longer measured and lengths will be extrapolated from the 30 measured fish. For some fish (like sticklebacks and shiner perch) were there were TWO DISTINCT size classes (juveniles and adults) 30 fish from each approximate age class were measured and fish counted but unmeasured were classified as *large* or *small* in the notes section of the csv files. 

This script is to clean the *dataset from 2020* of the eelgrass-associated fish community. The 2020 seines occured in July.

This script will:

1. Data checking 2020 seines and adjusting dataframe appropriately
2. Classify invertebrates and vertebrates so that only fish can be filtered
3. Convert unmeasured fish to *length measurements*
4. Convert length data to *biomass*
5. Convert date to julian day and add year column
6. Change site_code to new naming scheme that includes two column codes


# Libraries

```{r libraries, include = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(readxl)

```

# Data input

```{r data}
seine <- read.csv("Data/Fish_Seine/fish_beach_seine_2020_inprog-3_checked.csv") 

sp_names <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Abc823c8e-7be3-444b-a872-2e450ea3e85b"), stringsAsFactors = FALSE, header = TRUE)

site_names <- read.csv("Data/Site_universalnames_allyears_updated2022.csv")

fishLW <- read.csv("Data/fish_length_weight_conversions_cleaned_2021.csv", stringsAsFactors = FALSE, header = TRUE) 
```

# Step 1. Data checking

## Basic data checking
```{r}
seine <- seine %>%
  unite(bay_ID, bay_code:bay_sample)
names(seine)

unique(seine$bay_ID)
# Data entry mistake where there are multiple dates associated with one site/sampling event
unique(seine[,c("bay_ID","place_name", "date", "YYYYMMDD", "start_time", "end_time", "slope", "tide_height", "tide_time")])
# things to fix: 
# - remove lines NAs and one row with NFEI_
seine2 <- seine %>%
  filter(!is.na(YYYYMMDD)) %>%
  dplyr::select(-c(data_checking_FL:same_unmeas_sm.))

# check
unique(seine2[,c("bay_ID", "place_name", "date", "YYYYMMDD", "start_time", "end_time", "slope", "tide_height", "tide_time")])
```

```{r}
glimpse(seine2)
unique(seine2$sp_code) # 39 species including crabs

fish <- seine2 %>%
  filter(taxon == "vertebrate")
  

anti_join(fish, fishLW, by = "sp_code") %>%
  select(sp_code) %>%
  distinct()
# species codes in seines not in fish LW are: 
#- UNARTE
#- SCULSCA
#- UNSCUL
#- UNROCK

fish$sp_code <- as.factor(fish$sp_code)
# change SCULSCA scalyhead sculpin, should be classified with Artedius sp.
# This is because fish within the Artedius genus are hard to classify to species level. 
# Sometimes 
levels(fish$sp_code)[levels(fish$sp_code)=="SCULSCA"] <- "UNARTE"


# change sp_code columns back to characters
fish$sp_code <- as.character(fish$sp_code)

# check
unique(fish$sp_code) #Now 31 species
anti_join(fish, sp_names, by = c("sp_code" = "SpCode")) # nohthing missing all good

# Note: 
nonmatch1 <- anti_join(fish, fishLW, by = c("sp_code" = "sp_code"))
unique(nonmatch1$sp_code) # There a few more fish w/o length-weight relationship will be dealt with below. 
```

## Seperate size classes
In 2020 there were fish--like shiner perch that included two different size classes (adults and juveniles). Therefore when there were more than 30 individuals of those species 30 of *each* size class were measured creating a bimodial distribution. When extrapolating lengths to the unmeasured fish it is important to use probabilities from the correct distribution. 
# Use the notes section to do this 
```{r}
unique(fish$notes) #
names(fish) # has unmeasured_sm column
fish %>%
  filter(unmeasured_sm != "NA")
# the species where there are small subsets are SHIN 3spine and shiner perch
# REFU shiner
# NATZ 3spine and shiner
```
# Size distributions
For the two species that were subset into large and small subset what are their size distributions/frequency?
Is there a break between the "juvenile" and "adult" size classes? 
Three-spine Stickleback and Shiner Perch
```{r}
# subset data by species
shiner <- fish %>%
  filter(sp_code == "PERCHSH")

stickle <- fish %>%
  filter(sp_code == "STICK3")
```
Graph the size distributions of the measured fish across all sites
```{r}
shiner %>%
  ggplot() +
  geom_histogram(mapping = aes(x = fork_length), stat = "count") +
  geom_vline(aes(xintercept = 85, color = "red"))

#bimodal graph-- looks like the distributional cut off is close to 85 mm for shiner perch

stickle %>%
  #filter(site == "Salt Lake Bay") %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = fork_length), stat = "count") +
  geom_vline(aes(xintercept = 40, color = "red"))

# bimodial again, looks like 40 mm and below is the small cut off for three spine stickleback 
```

#Step 3. Unmeasured -> measured
This code comes from WR original script (Eelgrass_biometrics_data_analysis, L 325)

```{r}
#' From the raw data this script will export a file summarized by site ready for analysis. 
#' The goal will be to summaries this data at the site level. Summaries will occur in a few different
#' ways to prep the data for different types on analysis. There is 1 file generated here. 


#' All we need is the beach seine data and the fish length-mass conversions. 
#' make sure columns have appropriate character type

fish$unmeasured <- as.numeric(fish$unmeasured)
fish$unmeasured_sm <- as.numeric(fish$unmeasured_sm)
fish$fork_length <- as.numeric(fish$fork_length)

fish$abundance <- as.numeric(ifelse(is.na(fish$fork_length), ifelse(is.na(fish$unmeasured), paste(fish$unmeasured_sm), paste(fish$unmeasured)), 1))

df2 <- fish %>% # how many individual fish by site and species?
  dplyr::summarise(total = sum(abundance)) # 17442

df3 <- fish %>%
  group_by(bay_ID, sp_code) %>%
  dplyr::summarise(total = sum(abundance))
```
## Separate measured and unmeasured fish
### Loop for subset species
First run a similar loop as below to calculate estimated fork lengths for the small and large subset species. 
These include (for 2020) ONLY PERCHSH and 3SPINE at ONLY these sites: Shinaku, Refugion, Natzuhini, Nossuk
### 1. List of subset sites
```{r}
shiner %>%
  filter(is.na(fork_length))

stickle %>%
  filter(is.na(fork_length))
  
site_shiner <- c("SHIN_A", "NATZ_C")
site_stick <- c("SHIN_A", "REFU_A", "NATZ_C","NOSK_E", "NFEI_A")
```

### 2. Need to subset data by species and sites were they are subsampled
```{r}
fish.stick <- fish %>%
  filter(sp_code == "STICK3")

fish.shiner <- fish %>%
  filter(sp_code == "PERCHSH") 

```

### 3. subset by the sites where those species were subsampled
```{r}
fish.stick.sub <- subset(fish.stick, bay_ID %in% site_stick)
fish.shiner.sub <- subset(fish.shiner, bay_ID %in% site_shiner)
```

### 4. cycle through loops (1 for different species)
Sticklebacks first: based on distributions of known fork length we set a cut off at 40 mm 
```{r}
# Need dataframes by species, size, and measured v. unmeasured
sm.m.st <- fish.stick.sub %>%
  filter(fork_length != "NA") %>%
  filter(fork_length <= 40)
lg.m.st <- fish.stick.sub %>%
  filter(fork_length != "NA") %>%
  filter(fork_length > 40)
sm.um.st <- fish.stick.sub %>%
  filter(is.na(fork_length)) %>%
  filter(unmeasured_sm != "NA")
lg.um.st <- fish.stick.sub %>%
  filter(is.na(fork_length)) %>%
  filter(unmeasured != "NA")

```

Loop for stickleback
Loop for each size class to approximate measurements for unmeasured fish
```{r}
lg_stick <- data.frame() # empty dataframe to fill with for loop

for(s in lg.um.st$bay_ID){ # cycle through unique sites (only one seine per site)
  m <- lg.m.st %>% # subset measured data by interation EventID
    filter(bay_ID == s)
  u <- lg.um.st %>% # subset unmeasured data by iteration EventID
      filter(bay_ID == s)
  for(i in unique(fish.stick.sub$sp_code)){ # cycle through species that are in UNMEASURED data
    samp <- m %>% # create sample from which to make distrubution
      filter(sp_code == i)
    unmeas <- u %>% # isolate unmeasured fish
      filter(sp_code == i)
    unmeas <- (as.numeric(sum(unmeas$unmeasured))) # save unmeasured value
    dat.temp1 <- data.frame(size = as.numeric(samp$fork_length))
    dat.temp2 <- dat.temp1 %>% 
      group_by(size) %>% 
      dplyr::summarize(count = n())
    dat.temp2$prob <- (dat.temp2$count/sum(dat.temp2$count))
    dat.temp2$x <- as.numeric(dat.temp2$size)
    fx <- function(n){ # function derived from limits and probabilities of above
    sample(x = (dat.temp2$x), size = n, replace = TRUE, prob = dat.temp2$prob)
    }
    dat.temp3 <- data.frame(bay_ID = s, sp_code = i, fork_length = fx(unmeas))
    lg_stick <- rbind(lg_stick, dat.temp3) # append iteration to full data
    lg_stick[] <- lapply(lg_stick, as.character)
  }
} 

```

RE DO FOR THE SMALL SUBSET - sticklebacks
```{r}
sm_stick <- data.frame() # empty dataframe to fill with for loop

for(s in sm.um.st$bay_ID){ # cycle through unique sites (only one seine per site)
  m <- sm.m.st %>% # subset measured data by interation EventID
    filter(bay_ID == s)
  u <- sm.um.st %>% # subset unmeasured data by iteration EventID
      filter(bay_ID == s)
  for(i in unique(fish.stick.sub$sp_code)){ # cycle through species that are in UNMEASURED data
    samp <- m %>% # create sample from which to make distrubution
      filter(sp_code == i)
    unmeas <- u %>% # isolate unmeasured fish
      filter(sp_code == i)
    unmeas <- (as.numeric(sum(unmeas$unmeasured_sm))) # save unmeasured value
    dat.temp1 <- data.frame(size = as.numeric(samp$fork_length))
    dat.temp2 <- dat.temp1 %>% 
      group_by(size) %>% 
      dplyr::summarize(count = n())
    dat.temp2$prob <- (dat.temp2$count/sum(dat.temp2$count))
    dat.temp2$x <- as.numeric(dat.temp2$size)
    fx <- function(n){ # function derived from limits and probabilities of above
    sample(x = (dat.temp2$x), size = n, replace = TRUE, prob = dat.temp2$prob)
    }
    dat.temp3 <- data.frame(bay_ID = s, sp_code = i, fork_length = fx(unmeas))
    sm_stick <- rbind(sm_stick, dat.temp3) # append iteration to full data
    sm_stick[] <- lapply(sm_stick, as.character)
  }
} # this is returning an error in regards to prob, but i am not sure what its talking about
m_stick <- bind_rows(sm_stick, lg_stick)
```

Shiner perch: based on distributions of known fork length we set a cut off at 85 mm 
```{r}
# Need for data frames
sm.m.sh <- fish.shiner.sub %>%
  filter(fork_length != "NA") %>%
  filter(fork_length <= 85)
lg.m.sh <- fish.shiner.sub %>%
  filter(fork_length != "NA") %>%
  filter(fork_length > 85)
sm.um.sh <- fish.shiner.sub %>%
  filter(is.na(fork_length)) %>%
  filter(unmeasured_sm != "NA")
lg.um.sh <- fish.shiner.sub %>%
  filter(is.na(fork_length)) %>%
  filter(unmeasured != "NA")

```
Loop for shiner perch
```{r}
lg_shiner <- data.frame() # empty dataframe to fill with for loop

for(s in unique(fish.shiner.lg$bay_ID)){ # cycle through unique sites (only one seine per site)
  m <- lg.m.sh %>% # subset measured data by interation EventID
    filter(bay_ID == s)
  u <- lg.um.sh %>% # subset unmeasured data by iteration EventID
      filter(bay_ID == s)
  for(i in "PERCHSH"){ # cycle through species that are in UNMEASURED data
    samp <- m %>% # create sample from which to make distrubution
      filter(sp_code == i)
    unmeas <- u %>% # isolate unmeasured fish
      filter(sp_code == i)
    unmeas <- (as.numeric(sum(unmeas$unmeasured))) # save unmeasured value
    dat.temp1 <- data.frame(size = as.numeric(samp$fork_length))
    dat.temp2 <- dat.temp1 %>% 
      group_by(size) %>% 
      dplyr::summarize(count = n())
    dat.temp2$prob <- (dat.temp2$count/sum(dat.temp2$count))
    dat.temp2$x <- as.numeric(dat.temp2$size)
    fx <- function(n){ # function derived from limits and probabilities of above
    sample(x = (dat.temp2$x), size = n, replace = TRUE, prob = dat.temp2$prob)
    }
    dat.temp3 <- data.frame(bay_ID = s, sp_code = i, fork_length = fx(unmeas))
    lg_shiner <- rbind(lg_shiner, dat.temp3) # append iteration to full data
    lg_shiner[] <- lapply(lg_shiner, as.character)
  }
} 
# RE DO FOR THE SMALL SUBSET
sm_shiner <- data.frame() # empty dataframe to fill with for loop

for(s in unique(fish.shiner.sm$bay_ID)){ # cycle through unique sites (only one seine per site)
  m <- sm.m.sh %>% # subset measured data by interation EventID
    filter(bay_ID == s)
  u <- sm.um.sh %>% # subset unmeasured data by iteration EventID
      filter(bay_ID == s)
  for(i in "PERCHSH"){ # cycle through species that are in UNMEASURED data
    samp <- m %>% # create sample from which to make distrubution
      filter(sp_code == i)
    unmeas <- u %>% # isolate unmeasured fish
      filter(sp_code == i)
    unmeas <- (as.numeric(sum(unmeas$unmeasured_sm))) # save unmeasured value
    dat.temp1 <- data.frame(size = as.numeric(samp$fork_length))
    dat.temp2 <- dat.temp1 %>% 
      group_by(size) %>% 
      dplyr::summarize(count = n())
    dat.temp2$prob <- (dat.temp2$count/sum(dat.temp2$count))
    dat.temp2$x <- as.numeric(dat.temp2$size)
    fx <- function(n){ # function derived from limits and probabilities of above
    sample(x = (dat.temp2$x), size = n, replace = TRUE, prob = dat.temp2$prob)
    }
    dat.temp3 <- data.frame(bay_ID = s, sp_code = i, fork_length = fx(unmeas))
    sm_shiner <- rbind(sm_shiner, dat.temp3) # append iteration to full data
    sm_shiner[] <- lapply(sm_shiner, as.character)
  }
} 

m_shiner <- bind_rows(lg_shiner, sm_shiner)
```


Subset the rest of the data into m and um 
```{r}
# theres one fish (POATUBE) at NATZ_C that has one measured instances and one unmeasured, this makes the probabilities difficult so lets just list the one unmeasured as the measured length
m.poa <- fish %>%
  filter(bay_ID == "NATZ_C") %>%
  filter(sp_code == "POATUBE") %>%
  filter(fork_length != "NA")

fish.fixed <- fish %>%
  mutate(fork_length = ifelse(sp_code == "POATUBE" & bay_ID == "NATZ_C" & is.na(fork_length), paste(m.poa$fork_length), fork_length)) %>%
  mutate(unmeasured = ifelse(sp_code == "POATUBE" & bay_ID == "NATZ_C", NA, unmeasured))

# now seperate into measured and unmeasured
fish.m <- fish.fixed %>%
  filter(fork_length != "NA")

fish.um <- fish.fixed %>%
  filter(sp_code != "PERCHSH" & sp_code != "STICK3") %>%
  filter(is.na(fork_length)) #dont include any sites where there were um shiner/stick - already done in the loop above

```

## Assign lengths to unmeasured fish for other species

When beach seining we only measured the first 30 individuals of a species, and counted the rest. We can use the measured fishes to create a distribution from which we
can assign lengths to the unmeasured fishes.

Assign lengths to unmeasured fish based on sampled distribution. This assignment should happen at the EventID level. i.e. use the distribution of fishes at a EventID(seine instance) to assign unmeasured fishes at that EventID. 

We will assume the least and just use the sampled proportions to assign lenghts to unmeasured fish. Exclued fishes that do not have a measured counterpart. 
These were instances when we tossed a fish without knowing what it was other than it was a sculpin thing


Figure out which species at sites that there is not a measured conterpart 
```{r}
x <- fish.um %>%
  group_by(bay_ID, sp_code) %>%
  anti_join(fish.m, by = c("bay_ID", "sp_code")) # no instances of unmeasured fishes without a measured counterpart!
x
# want to remove these unmeasured fish from unmeasured df, in 2020 dont need to do this
#fish.um.redu <- fish.um %>%
#  group_by(site, sp_code) %>%
#  anti_join(x, by = c("bay_ID", "sp_code"))
# for 2020 just rename fish.um

fish.um.redu <- fish.um
fish.um.redu$bay_ID <- as.factor(fish.um.redu$bay_ID)
# Now calculate a fork_length/total_length for each unmeasured fish based on the 
# distribution of the existing fish BY site (one site per seine).
```

## Run loop to fill unmeasured fish length to measured
```{r}
#i <- "PRICKSN" # for testing to make sure loop is working
#s <- "SHIN_A"
w <- data.frame() # empty dataframe to fill with for loop

for(s in unique(fish.um.redu$bay_ID)){ # cycle through unique sites (only one seine per site)
  m <- fish.m %>% # subset measured data by interation EventID
    filter(bay_ID == s)
  u <- fish.um.redu %>% # subset unmeasured data by iteration EventID
      filter(bay_ID == s)
  for(i in unique(u$sp_code)){ # cycle through species that are in UNMEASURED data
    samp <- m %>% # create sample from which to make distrubution
      filter(sp_code == i)
    unmeas <- u %>% # isolate unmeasured fish
      filter(sp_code == i)
    unmeas <- (as.numeric(sum(unmeas$unmeasured))) # save unmeasured value
    dat.temp1 <- data.frame(size = as.numeric(samp$fork_length))
    dat.temp2 <- dat.temp1 %>% 
      group_by(size) %>% 
      dplyr::summarize(count = n())
    dat.temp2$prob <- (dat.temp2$count/sum(dat.temp2$count))
    dat.temp2$x <- as.numeric(dat.temp2$size)
    fx <- function(n){ # function derived from limits and probabilities of above
    sample(x = (dat.temp2$x), size = n, replace = TRUE, prob = dat.temp2$prob)
    }
    dat.temp3 <- data.frame(bay_ID = s, sp_code = i, fork_length = fx(unmeas))
    w <- rbind(w, dat.temp3) # append iteration to full data
    w[] <- lapply(w, as.character)
  }
} 
# Bind the now measured fish with fish from above (stick and shiner)
w <- bind_rows(w, m_shiner, m_stick)
```
## Loop check/add site info
Append assigned lengths to master data with all site level data

```{r}
# Extract site and sp data
fish.site <- unique(fish[,c("bay_ID", "date", "YYYYMMDD", "start_time", "end_time", "slope", "tide_height", "tide_time")])

fish$species_common <- trimws(fish$species_common, "r") # remove hanging white space on right side
fish.sp <- unique(fish[,c("species_common", "species_scientific", "sp_code", "taxon")])
# to make sure it joins with duplicating we need to change the species_scientific and common for UNARTE
fish.sp <- fish.sp %>%
  mutate(species_scientific = ifelse(sp_code == "UNARTE" & species_scientific != "Artedius spp.", "Artedius spp.", species_scientific)) %>%
  mutate(species_common = ifelse(sp_code == "UNARTE" & species_common != "Juvenile Artedius", "Juvenile Artedius", species_common)) %>%
  distinct()

## Merge with loop output, the lengths of unmeasured fishes ##
d.info1 <- left_join(w, fish.site, by = "bay_ID")

## Add species detail ##
d.info <- merge(d.info1, fish.sp, by = "sp_code")

test <- anti_join(d.info, w, by = "sp_code")
test # none that didn't join
## Merge with original measured fishes ##
d.info$fork_length <- as.numeric(d.info$fork_length)
fish.m$fork_length <- as.numeric(fish.m$fork_length)
fish.all <- fish.m %>%
  dplyr::select(bay_ID, date, YYYYMMDD, start_time, end_time, slope, tide_height,
                tide_time, species_common, species_scientific, sp_code, fork_length, 
                taxon, notes) %>%
  bind_rows(d.info)

# Do some basic checks to make sure the the conversion from unmeasured to measured worked correctly
unique(fish.all$bay_ID) # 6 sites, correct
unique(fish.all$sp_code) # 31 levels of fish species
unique(fish$sp_code) # 31 levels of fish in fish

# count total number of fish in original dataframe
names(fish)
#fish$abundance <- as.numeric(ifelse(is.na(fish$fork_length), paste(fish$unmeasured), 1)) 

fish %>%
  dplyr::summarise(counts = sum(abundance)) # 17442

df4 <- fish %>%
  group_by(bay_ID, sp_code) %>%
  dplyr::summarise(total = sum(abundance))

# does this match the fish.all dataframe?
fish.all$abundance <- as.numeric(ifelse(is.na(fish.all$fork_length), paste(fish.all$unmeasured), 1)) 

fish.all %>%
  dplyr::summarise(counts = sum(abundance)) #17442, it matches!

df5 <- fish.all %>%
  group_by(bay_ID, sp_code) %>%
  dplyr::summarise(counts = sum(abundance))

# check w/ previous numbers (df4) If there are issues with more individuals of e/a species after the forloop to 
# convert unmeasured fish to measured fish make sure to check the extracted site/date info. When you merge the loop
# output with the site info if there are duplicated sites, it will duplicate the number of fish--artificially 
# inflating the number of fish at ea site. 
merge(df5, df4)
```

# Step 4. Convert length to biomass
## Calculate Biomass
Using the length-weight conversion values individual lengths will be converted to biomass. Coefficients are in cm*g so fork lengths will need to be converted to cm from mm.

First the L-W conversion data will need to be prepped for use. This will include defineing one a value (collapse the a TL column), define sp_code for new species, and filter out estimates that were made from SL (standard length) and fish calssified as UNFISH or unidentified fish, and then summarise by taking the average a and b value for each species.


## Put parameters into one column 
```{r}
fishLW$a_cm.g <- ifelse(is.na(fishLW$a_cm.g), fishLW$aTL_cm.g, fishLW$a_cm.g)

## define sp_code for new species ##
# Shorthorn sculpin #
fishLW$sp_code <- ifelse(fishLW$species_common == "Shorthorn sculpin", paste("SCULSHRN"), fishLW$sp_code)

# Longhorn sculpin #
fishLW$sp_code <- ifelse(fishLW$species_common == "Longhorn sculpin", paste("SCULLHRN"), fishLW$sp_code)

# Rockfish general #
fishLW$sp_code <- ifelse(fishLW$species_common == "Rockfish", paste("ROCKSEB"), fishLW$sp_code)

## Filter and summarise ##
LW <- fishLW %>%
  filter(Type != "SL") %>% 
  filter(sp_code != "UNFISH") %>% 
  group_by(sp_code) %>% 
  dplyr::summarise(a = mean(a_cm.g),
            b = mean(b_cm.g))
```

We know that we do not have estimates for a and b estiamtes for all the species so we need to apply other values to other species. First identify what species we need to have stand ins for:

1. "UNSCUL" - Unidentified sculpins will use the average a and b values for all sculpins species.
2. "UNFLAT" - Unidentified flatfish will use the average a and b values for all flatfish species.
3. "UNGREEN" - Unidentified greenling will use the average a and b values for all greenling speices.
4. "UNMYOXO" - Unidentified Myoxocephalus sp. will use the the average of a and b values of sampled members of the Myoxocephalus genus.
5. "UNGUNN" - Unidentified gunnel will use average a and b values from the sampled gunnel species
6. "UNARTE" - Unidentified Aretedius sp. will use the average a nad b values of sampled members of the Artedeius genus.
7. "UNSNAIL" - Unidentified Liparis sp. will use the average a nad b values of sampled members of the Liparidae genus.

## Unique species in master data
```{r}
sp <- data.frame(sp_code = unique(fish.all$sp_code))

## Species that need stand ins ##
nomatch <- anti_join(sp, LW, by = "sp_code")
nonmatch

## Summarising stand ins ##
# Sculpins
scul <- data.frame(sp_code = "UNSCUL", fishLW %>% 
        filter(sp_code == "SCULBUF" | sp_code == "SCULGRT" | sp_code == "SCULGRU" |
                 sp_code == "SCULLHRN" | sp_code == "SCULMAN" |sp_code == "SCULPAD" | 
                 sp_code == "SCULPSTG" | sp_code == "SCULSAIL" | sp_code == "SCULSHRN" | 
                 sp_code =="SCULSILV" | sp_code == "SCULSMO" | sp_code == "SCULTIDE") %>% 
        summarise(a = mean(a_cm.g),
                  b = mean(b_cm.g)))

# Salmon
sal <- data.frame(sp_code = "SALSOCK", fishLW %>% 
        filter(sp_code == "SALCHIN" | sp_code == "SALCOHO" | sp_code == "SALPINK" |
                 sp_code == "SALCHUM") %>% 
        summarise(a = mean(a_cm.g),
                  b = mean(b_cm.g)))
# Flats
flat <- data.frame(sp_code = "UNFLAT", fishLW %>% 
                     filter(sp_code == "SDABPAC" | sp_code == "SDABSPKL" | sp_code == "SOLEBUT" |
                              sp_code == "SOLECO" | sp_code == "SOLEENG" |sp_code == "SOLEROC" | 
                              sp_code == "FLOUNST") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

# Greenlings
grns <- data.frame(sp_code = "UNGREEN", fishLW %>% 
                     filter(sp_code == "GREENKEL" | sp_code == "GREENMAS" | sp_code == "GREENWHI" |
                              sp_code == "GREENPAI") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

# Myoxocephalus
myoxo <- data.frame(sp_code = "UNMYOXO", fishLW %>% 
                      filter(sp_code == "SCULGRT" | sp_code == "SCULLHRN" | sp_code == "SCULSHRN") %>% 
                      summarise(a = mean(a_cm.g),
                                b = mean(b_cm.g)))

# Gunnels
gunn <- data.frame(sp_code = "UNGUNN", fishLW %>% 
                     filter(sp_code == "GUNNCRE" | sp_code == "GUNNPEN") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

# Artedius
arte <- data.frame(sp_code = "UNARTE", fishLW %>% 
                     filter(sp_code == "SCULPAD" | sp_code == "SCULSMO") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))
# Rockfish
rock <- data.frame(sp_code = "UNROCK", fishLW %>% 
                     filter(sp_code == "ROCKSEB" | sp_code == "ROCKSLG"
                            | sp_code == "ROCKCOP" | sp_code == "ROCKQUI"
                            | sp_code == "ROCKBLA" | sp_code == "ROCKBRO") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))

#Snailfish
snail <- data.frame(sp_code = "UNSNAILF", fishLW %>% 
                     filter(sp_code == "SNAILFR" | sp_code == "SNAILFTI"
                            | sp_code == "SNAILFSH" | sp_code == "SNAILFRT") %>% 
                     summarise(a = mean(a_cm.g),
                               b = mean(b_cm.g)))
## Master a b lsit ##
LW.master <- rbind.data.frame(LW, scul, flat, grns, myoxo, gunn, arte, sal, rock, snail)

## Filter out UNFISH from fish.all ##
fish.all <- fish.all %>% 
  filter(sp_code != "UNFISH")

## Check that everything is cool ##
unique(anti_join(fish.all, LW.master, by = "sp_code")$sp_code) # no unmatched sp_codes! awesome
```

```{r}
# Finally join the a and b values from above with the master data and calcualte biomass.
## Merge a and b values ##
fish.all.m <- merge(fish.all, LW.master, by = "sp_code")

## Convert mm to cm ##
fish.all.m$fork_length_cm <- fish.all.m$fork_length / 10

## Calculate mass in g, a*FL^b ##
fish.all.m$mass_g <- (fish.all.m$a * fish.all.m$fork_length_cm^fish.all.m$b)

## some plots, cuz ya know ##
hist(fish.all.m$mass_g)
hist(log(fish.all.m$mass_g))
range(fish.all.m$mass_g)

## Clean up ##
fish.all.m$taxon[is.na(fish.all.m$taxon)] <- "Vertebrata"

# remove unused rows and readd in tide height
# extra tide height from original df
tide_h <- fish %>%
  group_by(bay_ID) %>%
  dplyr::select(bay_ID, tide_height) %>%
  distinct(tide_height)

fish.all.m <- fish.all.m %>%
  left_join(tide_h, by = "bay_ID") %>%
  mutate(tide_height = tide_height.x) %>%
  dplyr::select(-c(abundance, tide_height.y, tide_height.x))
  

## Checking ##
fish.site.g <- fish.all.m %>% 
  group_by(bay_ID) %>% 
  dplyr::summarise(site_mass_g = sum(mass_g))

fish.site.g$mass_kg <- fish.site.g$site_mass_g / 1000

boxplot(fish.site.g$site_mass_g)

```

# Step 5. Convert date to JD and add year column
```{r}
fish.all.m <- fish.all.m %>% 
  mutate(date = ymd(YYYYMMDD)) %>% # convert to date format
  mutate(julian = yday(date)) %>% # make the date a julian day
  mutate(year = year(date)) # make a new column with just year
```

# Step 6. Convert site names to 2-column code system
```{r}
site_names <- site_names %>%
  unite(bay_ID, bay_code:bay_sample)

names(site_names)

fish.all.m2 <- fish.all.m %>% 
 left_join(site_names, by = "bay_ID") %>% # this needs to be changed for e/a YEAR. 
 select(-c(site_2017, site_2018, site_2019, freshwater, general_description, sediment_description, siteID_NOAA, study, notes,
           fork_length_cm, YYYYMMDD, a, b)) %>%
  separate(bay_ID, into = c("bay_code", "bay_sample"))

names(fish.all.m2)
# place name gives the physical location, bay_code and bay_sample together give you each
# unique site that was sampled. 

```

Do some basic checks before writing out the csv
```{r}
unique(fish.all.m$bay_ID) # 6 sites!
unique(fish.all.m$sp_code) # 31 species!
range(fish.all.m$mass_g) # makes sense there were some tiny fish. 
```


# Export
```{r}
#write.csv(fish.all.m2, "Data/Fish_seine/fish_mass_2020_derived.csv", row.names = FALSE)
```


